---
title: "02 Cultivation score weights"
output: html_notebook
---

# Goal

Determine the extent to which each of the data elements related to the PG cultivation score (see [Ben Porter's Donor Cultivation Checklist](https://www.case.org/currents/x74757)) have impacted campaign giving.

# Setup and datafile

```{r setup, message = FALSE, warning = FALSE}
library(tidyverse)
```

The data file is generated with [this code](https://github.com/phively/ksm-models/blob/master/pg-cultivation-score-fy18/code/generate-data.R), adapted from the import data step of [01 Initial exploration.Rmd
](https://github.com/phively/ksm-models/blob/master/pg-cultivation-score-fy18/01%20Initial%20exploration.Rmd).

```{r}
source('code/generate-data.R')
```

The following 12 items were able to be extracted from the database:

  * Alum or spouse of alum
  * Age
  * Chicago home address
  * Visited by prospect manager in last 2 years
  * 5+ total visits
  * Deep engagement (multiple degrees, parent, season tickets, etc.)
  * High-level annual giving ($25K+)
  * High-level advisory board participation
  * Previous major gift ($250K+)
  * Meeting with president
  * Open proposal
  * Consistent donor (10+ years of giving, including 1+ of last 3)

Each counts as one point toward the cultivation score, which ranges from 0 to `r max(pool$CULTIVATION_SCORE) %>% I()`.

# Data exploration

