---
title: "04 KSM campaign model"
output:
  html_notebook:
    code_folding: hide
    toc: TRUE
    toc_float:
      collapsed: FALSE
---

# Goal

Supplement the previous models with Kellogg-specific datapoints previously used for the [AF $10K model](https://github.com/phively/ksm-models/tree/master/af-10k-fy17), assess how well the PG checklist applies to Kellogg donors, and build a basic campaign prioritization model using all variables.

# Setup

```{r setup, message = FALSE, warning = FALSE}
library(tidyverse)
library(gridExtra)
library(splines)
library(foreach)
library(lubridate)
library(wranglR)

# Functions adapted from previous analysis steps
source('code/functions.R')

# Visualization functions adapted fron previous analysis steps
source('code/functions_viz.R')
```

## Import data

The PG checklist data file is generated with [this code](https://github.com/phively/ksm-models/blob/master/pg-cultivation-score-fy18/code/generate-data.R), while the KSM data is derived from the 1/9/2018 version of [this view](https://github.com/phively/ksm-models/blob/master/af-10k-fy17/data/ksm-af-10k-data.sql), transformed by [a script](https://github.com/phively/ksm-models/blob/master/pg-cultivation-score-fy18/code/generate-af10k-data.R).

```{r, warning = FALSE}
date <- '2018-09-07'

# Load PG data
filepath <- paste0('data/', date, ' PG scores for all active prospects.xlsx')
source('code/generate-data.R')
source('code/generate-modeling-data.R')
# Preserve ID numbers
mdat <- pool %>% select(ID_NUMBER) %>% cbind(mdat)

# Load AF 10K data
filepath <- paste0('data/', date, ' AF 10K model data.csv')
source('code/generate-af10k-data.R')
dat10k <- parse_af_data(filepath)
```

# KSM and NU model comparison

## Campaign vs. PG checklist

Does fitting the PG checklist on all prospects, versus Kellogg prospects only, show some difference? How does the model change when fitting just KSM prospects, as opposed to the entire prospect pool?

```{r}
# Full dataset
chklm_full <- lm(
  log10plus1(CAMPAIGN_NEWGIFT_CMIT_CREDIT) ~
    ACTIVE_PROPOSALS +
    AGE +
    PM_VISIT_LAST_2_YRS +
    VISITS_5PLUS +
    AF_25K_GIFT +
    GAVE_IN_LAST_3_YRS +
    MG_250K_PLUS +
    PRESIDENT_VISIT +
    TRUSTEE_OR_ADVISORY_BOARD +
    Alumnus +
    DEEP_ENGAGEMENT +
    CHICAGO_HOME
  , data = mdat
)

# KSM prospects only
chklm_ksm <- lm(
  log10plus1(CAMPAIGN_NEWGIFT_CMIT_CREDIT) ~
    ACTIVE_PROPOSALS +
    AGE +
    PM_VISIT_LAST_2_YRS +
    VISITS_5PLUS +
    AF_25K_GIFT +
    GAVE_IN_LAST_3_YRS +
    MG_250K_PLUS +
    PRESIDENT_VISIT +
    TRUSTEE_OR_ADVISORY_BOARD +
    Alumnus +
    DEEP_ENGAGEMENT +
    CHICAGO_HOME
  , data = mdat %>% inner_join(dat10k %>% select(ID_NUMBER), by = c('ID_NUMBER', 'ID_NUMBER'))
)
```

The full model is fit on all `r chklm_full %>% model.frame() %>% nrow() %>% I()` observations, while the KSM model uses only `r chklm_ksm %>% model.frame() %>% nrow() %>% I()` of them.

```{r}
model_comparison <- full_join(
  data.frame(
    var = coef(chklm_full) %>% names
    , full.model = coef(chklm_full)
    , se.full = summary(chklm_full)$coefficients[, 2]
  )
  , data.frame(
    var = coef(chklm_ksm) %>% names
    , ksm.model = coef(chklm_ksm)
    , se.ksm = summary(chklm_ksm)$coefficients[, 2]
  )
  , by = c('var', 'var')
)
```

```{r, rows.print = 100}
model_comparison %>%
  mutate_if(is.numeric, function(x) round(x, 3))
```

```{r, fig.height = 6}
pcrit <- .05
critq <- qnorm(p = pcrit/2, lower.tail = FALSE)
model_comparison %>%
  gather('which', 'x', -(var)) %>%
  mutate(
    model = case_when(
      str_detect(which, '.*full.*') ~ 'full'
      , str_detect(which, '.*ksm.*') ~ 'ksm'
    ) %>% factor()
    , which = case_when(
      str_detect(which, '.*model.*') ~ 'beta'
      , TRUE ~ 'se'
    )
  ) %>%
  spread(which, x) %>%
  mutate(
    se.high = beta + se * critq
    , se.low = beta - se * critq
  ) %>%
  ggplot(aes(x = var, y = beta, color = model, group = model)) +
  geom_hline(yintercept = 0, color = 'darkgray') +
  geom_point(position = position_dodge(.5)) +
  geom_linerange(aes(ymin = se.low, ymax = se.high), position = position_dodge(.5)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = .5), axis.title.y = element_text(angle = 0, vjust = .5)) +
  labs(x = 'variable', y = bquote(hat(beta)))
```

The bars provide `r scales::percent(1 - pcrit) %>% I()` CIs for each coefficient based on the standard error. Not much here; there isn't an apparent difference for the variables of interest. How about the final campaign model?

## Interaction model

To be more rigorous, consider a "KSM prospect" indicator, with interaction effects on each of the explanatory variables.

```{r}
clm_ksm_interact <- lm(
  log10plus1(CAMPAIGN_NEWGIFT_CMIT_CREDIT) ~
    # KSM interaction
    KSM *
    (
    ns(NUMERIC_AGE, df = 5) + # Underlying variable to AGE indicator
    PM_VISIT_LAST_2_YRS +
    log10plus1(VISIT_COUNT) + # Underlying VISITS_5PLUS indicator
    AF_25K_GIFT +
    sqrt(YEARS_OF_GIVING) + # Underlying GAVE_IN_LAST_3_YRS
    ns(YEARS_OF_GIVING_LAST_3, df = 2) + # Underlying GAVE_IN_LAST_3_YRS
    MG_250K_PLUS + # Decided to leave as factor
    Alumnus +
    SEASON_TICKET_YEARS + # Deep Engagement component
    AFFINITY_SCORE +
    MG_PR_MODEL_DESC +
    1
    )
  , data = mdat %>%
      # People on the 10K list count as KSM prospects
      left_join(dat10k %>% select(ID_NUMBER) %>% mutate(KSM = 'Y')
                , by = c('ID_NUMBER', 'ID_NUMBER')) %>%
      # Replace NAs from the left join with N indicator
      replace_na(replace = list(KSM = 'N'))
)
```

```{r}
summary(clm_ksm_interact)
```

Note the NAs due to collinearity; this should be fixed.

```{r}
clm_ksm_interact <- update(
  clm_ksm_interact
  , formula = . ~ .
      - (ns(YEARS_OF_GIVING_LAST_3, df = 2) * KSM)
      + (YEARS_OF_GIVING_LAST_3 * KSM)
)
```

```{r}
summary(clm_ksm_interact)
```

```{r}
interact_model <- cbind(
  data.frame(var = summary(clm_ksm_interact)$coefficients %>% rownames())
  , beta = summary(clm_ksm_interact)$coefficients[, 1]
  , se = summary(clm_ksm_interact)$coefficients[, 2]
)
row.names(interact_model) <- NULL
```

```{r, fig.height = 6}
interact_model %>%
  filter(str_detect(var, 'KSMY')) %>%
  mutate(
    var = str_remove(var, ':KSMY') %>% as_factor()
    , se.high = beta + se * critq
    , se.low = beta - se * critq
    # SE bars cross 0 iff not significant at pcrit
    , sig = ifelse(se.high > 0 & se.low < 0, 'Not significant', 'Significant')
  ) %>%
  ggplot(aes(x = var, y = beta, color = sig)) +
  geom_hline(yintercept = 0, color = 'darkgray') +
  geom_point() +
  geom_linerange(aes(ymin = se.low, ymax = se.high)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = .5), axis.title.y = element_text(angle = 0, vjust = .5)) +
  labs(
    title = 'KSM vs. NU campaign model interaction coefficeints'
    , x = 'variable'
    , y = bquote(hat(beta))
    , color = bquote('Significance at p =' ~ .(pcrit) ~ 'level')
  )
```

Looks like there *is* a KSM main effect (KSMY is the KSM effect), plus there are differences in the age (less impact), total visit count (more impact), years of giving (including last 3, at the end), affinity score (slightly less impact), and PG prioritization model factors (less impact).

```{r}
interact_compare <- interact_model %>%
  filter(str_detect(var, 'NUMERIC_AGE|VISIT_COUNT|YEARS_OF_GIVING|AFFINITY|MG_PR')) %>%
  mutate(
    model = ifelse(str_detect(var, 'KSMY'), 'KSM', 'NU')
    , var = str_remove(var, ':KSMY')
  )
interact_compare <- cbind(
  interact_compare %>% filter(model == 'NU')
  , interact_compare %>% filter(model == 'KSM') %>% select(ksm.effect = beta, ksm.se = se)
) %>% mutate(
  ksm.effect = beta + ksm.effect
) %>% select(-model)
```

```{r, rows.print = 100}
interact_compare %>%
  mutate_if(is.numeric, function(x) round(x, 3))
```

I've defined "KSM effect" as the sum of the coefficient and interaction effect estimate, i.e.

$$ K_i = \hat{\beta_i} + \hat{\beta}_{i ~ \times KSM} $$

Plotting these variables:

```{r}
interact_compare %>%
  gather('which', 'x', -(var)) %>%
  mutate(
    model = case_when(
      !str_detect(which, 'ksm.*') ~ 'NU estimate'
      , TRUE ~ 'KSM effect'
    ) %>% as_factor()
    , which = case_when(
      str_detect(which, 'se') ~ 'se'
      , TRUE ~ 'beta'
    )
  ) %>%
  spread(which, x) %>%
  mutate(
    se.high = beta + se * critq
    , se.low = beta - se * critq
  ) %>%
  ggplot(aes(x = var, y = beta, color = model, group = model)) +
  geom_hline(yintercept = 0, color = 'darkgray') +
  geom_point(position = position_dodge(.5)) +
  geom_linerange(aes(ymin = se.low, ymax = se.high), position = position_dodge(.5)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = .5), axis.title.y = element_text(angle = 0, vjust = .5)) +
  labs(x = 'variable', y = bquote(hat(beta)), color = bquote(hat(beta) ~ 'or effect')) +
  scale_y_continuous(breaks = seq(-5, 5, by = .5))
```

This is interesting -- there does appear to be evidence that middle tier and top tier aren't as robust for KSM (coefficients closer to 0), but affinity score still looks pretty good. The practical effect of age coefficient differences for KSM/NU is relatively large, even changing signs for most of the splines. Total visit count and total years of giving is also more impactful for KSM than NU prospects.

## Summary of findings

  * KSM and NU constituents behave similarly when looking only at the PG checklist items.
  * Given the interaction model results, it probably makes sense to estimate age effects separately for the different units/programs.
  * There really are differences between KSM and NU campaign giving not already accounted for in the MG prioritization score. Note that this is **not** the same as saying that expected *future giving* (i.e. the purpose of the model) is different between the populations in some way not accounted for by the MG prioritzation score.
  * Affinity score is a marginally less important predictor for KSM than NU prospects. On the other hand, total visit count and total years of giving outperform for KSM compared to NU.
  * The previous point may suggest that giving engagement is *more* important for KSM prospects (and potentially even other graduate program prospects) than other NU prospects.

# KSM model variable selection
