---
title: "06 KSM model tuning"
output:
  html_notebook:
    code_folding: hide
    toc: TRUE
    toc_float:
      collapsed: FALSE
---

# Goal

Construct a final two-stage model based on my previous work, computing:

$$ E \left( \text{giving, donor | covariates} \right) = E \left(\text{giving | donor, covariates} \right) P \left(\text{donor | covariates} \right) $$

# Setup

```{r setup, message = FALSE, warning = FALSE}
library(tidyverse)
library(reshape2)
library(gridExtra)
library(splines)
library(lubridate)
library(wranglR)
library(Boruta)
library(foreach)
library(doParallel)
# Functions adapted from previous analysis steps
source('code/functions.R')
# Visualization functions adapted fron previous analysis steps
source('code/functions_viz.R')
# Set number of available CPU cores
registerDoParallel(10)
```
```{r}
# Parameters
train_fy <- 2016
filepath <- 'data/2018-11-30 point-in-time data.xlsx'
sheetname <- 'Select point_in_time_model'

# Import data
source('code/generate-pit-data.R')

# Run data generation functions
modeling.data <- generate_pit_data(filepath, sheetname) %>%
  generate_additional_predictors()
```
```{r}
# Create cross-validation set
folds = 10
reps = 5

# Withhold 10% of data as test set
xv <- KFoldXVal(modeling.data, k = 2, prop = .1, seed = 6988432)
holdoutdat <- modeling.data[xv[[1]], ]
traindat <- modeling.data[xv[[2]], ]
remove(xv)
```

# Probability model tuning

I'll perform a grid search to select reasonable spline parameters. To cut down on the number of combinations, similar variables will have the same number of spline dfs.

```{r}
# Splines grid search parameters
grid_params = list(
  giving = 2:3
  , dollars = 2:4
  , engagement = 2:3
  , recordyr = seq(3, 7, by = 2)
)
```
```{r}
print(grid_params)
```


This gives `r lapply(grid_params, length) %>% unlist() %>% prod() %>% I()` combinations.

```{r}
# Store timings
glm_timestamps <- list()
# Store model errors
glm_errs <- list()
# Seed for reproducibility
set.seed(229786270)

# Outer loop (repetitions)
for (rep in 1:reps) {
  # Status report 
  timestamp <- paste('+ Iteration', rep, 'beginning at:', Sys.time())
  print(timestamp)
  glm_timestamps <- c(glm_timestamps, timestamp)
  # Create cross-validation indices
  xv <- KFoldXVal(traindat, k = folds)
  # Middle loop (grid search)
  errs_out <- foreach(
    giving = grid_params$giving
    , .combine = c
    , .multicombine = TRUE
  ) %:% foreach(
    dollars = grid_params$dollars
    , .combine = c
    , .multicombine = TRUE
  ) %:% foreach(
    engagement = grid_params$engagement
    , .combine = c
    , .multicombine = TRUE
  ) %:% foreach(
    recordyr = grid_params$recordyr
    , .combine = c
    , .multicombine = TRUE
  ) %:% foreach(
  # Inner loop (parallel cross-validation)
    fold = 1:length(xv)
    , .combine = list
    , .multicombine = TRUE
    , .packages = c('dplyr', 'splines')
  ) %dopar% {
    # Fit temp model
    tmpmodel <- glm(
      rv.gave ~
      PROGRAM_GROUP +
      PREF_ADDR_TYPE_CODE +
      HOUSEHOLD_CONTINENT +
      BUS_IS_EMPLOYED +
      HAS_HOME_ADDR +
      HAS_HOME_PHONE +
      ns(YEARS_SINCE_FIRST_GIFT, df = giving) +
      ns(GIVING_FIRST_YEAR_CASH_AMT, df = giving) +
      ns(GIVING_MAX_PLEDGE_AMT, df = giving) +
      ns(GIVING_CASH_TOTAL, df = giving) +
      ns(GIVING_PLEDGE_TOTAL, df = giving) +
      ns(GIVING_CRU_TOTAL, df = giving) +
      ns(GIFTS_ALLOCS_SUPPORTED, df = giving) +
      ns(GIFTS_FYS_SUPPORTED, df = giving) +
      ns(GIFTS_CASH, df = giving) +
      ns(GIFTS_PLEDGES, df = giving) +
      ns(CASH_PFY1, df = dollars) +
      ns(CASH_PFY2, df = dollars) +
      ns(CASH_PFY3, df = dollars) +
      ns(CASH_PFY4, df = dollars) +
      ns(CASH_PFY5, df = dollars) +
      CRU_GIVING_SEGMENT +
      ns(EVALUATION_LOWER_BOUND, df = dollars) +
      ns(UOR_LOWER_BOUND, df = dollars) +
      ns(MONTHS_ASSIGNED, df = engagement) +
      ns(COMMITTEE_NU_DISTINCT, df = engagement) +
      ns(COMMITTEE_NU_YEARS, df = engagement) +
      ns(COMMITTEE_KSM_DISTINCT, df = engagement) +
      ns(EVENTS_PREV_3_FY, df = engagement) +
      ns(EVENTS_CFY, df = engagement) +
      ns(EVENTS_PFY1, df = engagement) +
      ns(ATHLETICS_TICKET_YEARS, df = engagement) +
      ns(YEARS_SINCE_ATHLETICS_TICKETS, df = engagement) +
      ns(RECORD_YR, df = recordyr) +
      ns(YEARS_SINCE_MAX_CASH_YR, df = giving) +
      GIVING_MAX_CASH_MO +
      KSM_PROSPECT +
      ns(VISITORS_5FY, df = engagement) +
      LOYAL_5_PCT_CASH +
      UPGRADE3_CASH +
      VELOCITY3_LIN_CASH +
      SPOUSE_ALUM
      , data = traindat[-xv[[fold]], ]
      , family = 'binomial'
    )
    # Prediction threshold
    theta1 <- sum(traindat$rv.gave[-xv[[fold]]] == 1) / nrow(traindat[-xv[[fold]], ])
    # Confusion matrix based on the withheld data
    tmpconfus <- conf_matrix(tmpmodel, newdata = traindat[xv[[fold]], ], rv = 'rv.gave', threshold = theta1)
    # Return results
    return(
      list(
        conf_matrix = tmpconfus$conf_matrix
        , conf_matrix_pct = tmpconfus$conf_matrix_pct
        , errors = data.frame(
          reps = rep
          , folds = fold
          , giving = giving
          , dollars = dollars
          , engagement = engagement
          , recordyr = recordyr
          , error = tmpconfus$error
          , precision = tmpconfus$precision
          , sensitivity = tmpconfus$sensitivity
          , F1_score = tmpconfus$F1_score
        )
      )
    )
  }
  # Write results to errors data frame
  glm_errs <- c(glm_errs, errs_out)
  # Status report
  timestamp <- paste(' -Iteration', rep, 'ending at:   ', Sys.time())
  print(timestamp)
  glm_timestamps <- c(glm_timestamps, timestamp)
}
```
```{r}
for(i in 1:length(glm_timestamps)) {print(glm_timestamps[[i]])}
```
```{r, include = FALSE}
# Code to create a nested list from an un-nested list
# Not needed now that the code above uses .multicombine = TRUE
# tmp <- list()
# for (i in 1:1800) {
# idx <- 3*i - 2
#   tmp[[i]] = list(
#     conf_matrix = glm_errs[[idx]]
#     , conf_matrix_pct = glm_errs[[idx + 1]]
#     , errors = glm_errs[[idx + 2]]
#   )
# }
```
```{r}
save(glm_errs, glm_timestamps, file = 'data/06_glm_xval.Rdata')
```

## Cross-validated errors

Examine the errors for each combination of spline dfs.

# Regression model