---
title: "07 KSM model refresh"
output:
  html_notebook:
    code_folding: hide
    toc: TRUE
    toc_float:
      collapsed: FALSE
---

# Goal

Create final prospect rankings based on expected future giving.

# Setup

```{r setup, message = FALSE, warning = FALSE}
library(tidyverse)
library(reshape2)
library(gridExtra)
library(splines)
library(lubridate)
library(wranglR)
library(Boruta)
library(foreach)
library(doParallel)
library(mgcv)
# Functions adapted from previous analysis steps
source('code/functions.R')
# Visualization functions adapted fron previous analysis steps
source('code/functions_viz.R')
# Set number of available CPU cores
registerDoParallel(10)
```
```{r}
# Load current data
# Parameters
train_fy <- 2018
filepath <- 'data/2019-01-09 point-in-time data.xlsx'
sheetname <- 'Select point_in_time_model_fy18'

# Import data
source('code/generate-pit-data.R')

# Run data generation functions
modeling.data <- generate_pit_data(filepath, sheetname) %>%
  generate_additional_predictors(future.data = TRUE)
```
```{r}
# Load historical data
train_fy <- 2016
historical.modeling <- generate_pit_data('data/2018-11-30 point-in-time data.xlsx', 'Select point_in_time_model') %>%
  generate_additional_predictors()
```
```{r}
# Load fitted models
load('data/06_gam_final.Rdata')
load('data/06_lm_final.Rdata')
```

# Generate new predictions

```{r}
preds_prb <- predict(gam_final, newdata = modeling.data, type = 'response') %>% as.numeric()
preds_amt <- predict(lm_final, newdata = modeling.data, type = 'response')
```

# Diagnostics

Compare the range of in-sample predictions to the new out-of-sample predictions.

```{r}
data.frame(
  data = 'insample'
  , prediction = gam_final$fitted.values
) %>% rbind(
  data.frame(
    data = 'outsample'
    , prediction = preds_prb
  )
) %>%
  ggplot(aes(x = prediction, fill = data)) +
  geom_histogram(bins = 20) +
  facet_wrap(. ~ data) +
  labs(title = 'Probability model fitted values')
```

```{r}
data.frame(
  data = 'insample'
  , prediction = predict(lm_final, newdata = historical.modeling)
) %>% rbind(
  data.frame(
    data = 'outsample'
    , prediction = preds_amt
  )
) %>%
  ggplot(aes(x = prediction, fill = data)) +
  geom_histogram(bins = 20) +
  facet_wrap(. ~ data) +
  labs(title = 'Gift amount model fitted values')
```

Looks reasonable to me.

Compare the FY16 predictions and FY18 results for individuals -- how much have individuals moved?

```{r}
data.frame(
  id = historical.modeling$ID_NUMBER %>% as.numeric()
  , p_insample = gam_final$fitted.values
) %>% inner_join(
  data.frame(
    id = modeling.data$ID_NUMBER %>% as.numeric()
    , p_outsample = preds_prb
  )
  , by = c('id')
) %>% mutate(
  change = p_insample - p_outsample
) %>%
  ggplot(aes(x = change)) + 
  geom_histogram(bins = 20, alpha = .5) +
  scale_x_continuous(breaks = seq(-1, 1, by = .2)) +
  scale_y_log10(breaks = 10^(0:6), labels = format(10^(0:6), scientific = FALSE)) +
  labs(title = 'Change in predicted probability of giving')
```

The near symmetry around 0 is reassuring.

```{r}
data.frame(
  id = historical.modeling$ID_NUMBER %>% as.numeric()
  , p_insample = predict(lm_final, newdata = historical.modeling)
) %>% inner_join(
  data.frame(
    id = modeling.data$ID_NUMBER %>% as.numeric()
    , p_outsample = preds_amt
  )
  , by = c('id')
) %>% mutate(
  change = p_insample - p_outsample
) %>%
  ggplot(aes(x = change)) + 
  geom_histogram(bins = 16, alpha = .5) +
  scale_y_log10(breaks = 10^(0:6), labels = format(10^(0:6), scientific = FALSE)) +
  labs(title = 'Change in predicted giving amounts')
```

Interestingly, more people are expected to downgrade than upgrade.

# Save results

```{r}
data.frame(
  est.giving = preds_amt
  , est.p = preds_prb
) %>% mutate(
  ev = est.giving * est.p
  , est.giving.bin = trunc(est.giving)
  , est.p.bin = trunc(est.p)
  , ev.bin = trunc(ev)
) %>% cbind(
  modeling.data
) %>% write.csv(
  file = 'data/2019-01-09 predictions.csv'
  , row.names = FALSE
)
```

