---
title: "10 Model effectiveness"
output:
  html_document:
    df_print: paged
    toc: yes
  html_notebook:
    code_folding: hide
    toc: yes
    toc_float:
      collapsed: no
---

# Goal

Now that we have about a year's worth of additional data I'd like to investigate the effectiveness or ROI of the model by looking at different metrics associated with outcomes during the out-of-sample period. Some ideas:

  * Giving (direct measure)
  * Visits (indirect measure)
  * Qualifications/disqualifications (indirect measure)
  
# Libraries

```{r setup, message = FALSE, warning = FALSE}
library(tidyverse)
library(ggplot2)
```

# Visit data

I created a SQL query ([roi-visit-outcomes.sql](https://github.com/phively/ksm-models/blob/master/pg-cultivation-score-fy18/code/roi-visit-outcomes.sql)) to look at people who were reached out to specifically for qualification purposes in the last two years.

```{r}
visits <- readxl::read_xlsx(
  path = 'data/2019-07-23 qualifications data.xlsx'
  , sheet = 'With'
)

# Add some helpful indicators
visits <- visits %>% mutate(
  # Only outreach since the model data year
  qual = CFY_QUALIFICATION + PFY1_QUALIFICATION > 0
  , qual_EXCL_VISIT = CFY_QUALIFICATION_EXCL_VISIT + PFY1_QUALIFICATION_EXCL_VISIT > 0
  , outreach = CFY_OUTREACH + PFY1_OUTREACH > 0
  , visit = CFY_VISITS + PFY1_VISITS + PFY2_VISITS > 0
  , scored = !is.na(PR_SCORE)
  , never_visited = is.na(FIRST_VISIT_FY)
  , fy19_first_visit = !never_visited & FIRST_VISIT_FY == 2019
)
```

```{r}
rbind(
  Total.contacted = visits %>% nrow()
  , With.PR.score = visits %>% filter(scored) %>% nrow()
  , Never.visited = visits %>% filter(never_visited) %>% nrow()
  , FY19.first.visit = visits %>% filter(fy19_first_visit) %>% nrow()
)
```

```{r}
rbind(
  With.outreach = visits %>% filter(scored & outreach) %>% nrow()
  , With.qual = visits %>% filter(scored & outreach) %>% filter(qual_EXCL_VISIT) %>% nrow()
  , With.outreach.visit = visits %>% filter(scored & outreach & visit) %>% nrow()
  , With.qual.visit = visits %>% filter(scored & qual_EXCL_VISIT & visit) %>% nrow()
)
```

```{r}
visits_first_potential <- visits %>% filter(scored) %>% filter(fy19_first_visit | never_visited)

rbind(
  With.outreach = visits_first_potential %>% filter(outreach) %>% nrow()
  , With.qual = visits_first_potential %>% filter(qual_EXCL_VISIT) %>% nrow()
  , With.outreach.visit = visits_first_potential %>% filter(outreach & visit) %>% nrow()
  , With.qual.visit = visits_first_potential %>% filter(qual_EXCL_VISIT & visit) %>% nrow()
)
```

This is more than enough first-time visits for an analysis. Begin by looking at qualification counts and visit outcomes by PR score.

```{r}
visits %>%
  # Only people with qualification attempts
  filter(scored & qual_EXCL_VISIT) %>%
  # Must have had non-visit qualification attempts last year
  filter(PFY1_QUALIFICATION_EXCL_VISIT > 0) %>%
  ggplot(aes(x = PR_SCORE)) +
  geom_point(aes(color = visit, y = as.numeric(visit))) +
  geom_density(color = 'darkgray') +
  geom_smooth(aes(y = as.numeric(visit)), method = 'loess') +
  labs(y = 'Visit proportion', title = 'FY18-19 qualification attempts leading to visits')
```


```{r}
visits_first_potential %>%
  # Only people with qualification attempts in the last couple years
  filter(qual_EXCL_VISIT | outreach) %>%
  ggplot(aes(x = PR_SCORE)) +
  geom_point(aes(color = fy19_first_visit, y = as.numeric(visit))) +
  geom_density(color = 'darkgray') +
  geom_smooth(aes(y = as.numeric(visit)), method = 'loess') +
  labs(y = 'First visit proportion', title = 'FY18-19 qualification attempts leading to FY19 first visit')
```

So that's somewhat promising.

# FY19 visit status predictions

Start by looking at people who took first visits this year based on last year's outreach.

```{r}
model <- visits_first_potential %>% glm(
  fy19_first_visit ~
    PR_SCORE +
    ID_SCORE +
    log10(EVALUATION_LOWER_BOUND + 1) +
    sqrt(PFY1_QUALIFICATION_EXCL_VISIT) +
    sqrt(PFY2_QUALIFICATION_EXCL_VISIT) +
    sqrt(PFY3_QUALIFICATION_EXCL_VISIT) +
    sqrt(PFY1_OUTREACH) + 
    sqrt(PFY2_OUTREACH) + 
    sqrt(PFY3_OUTREACH) +
    sqrt(PFY1_EMAIL) +
    sqrt(PFY2_EMAIL) +
    sqrt(PFY3_EMAIL) +
    sqrt(PFY1_PHONE) + 
    sqrt(PFY2_PHONE) +
    sqrt(PFY3_PHONE) +
    PROGRAM_GROUP
  , data = .
)

summary(model)
```

Or combining years:

```{r}
model_b <- visits_first_potential %>% mutate(
  ALLPFY_QUAL_EXCL_VISIT = PFY1_QUALIFICATION_EXCL_VISIT + PFY2_QUALIFICATION_EXCL_VISIT + PFY3_QUALIFICATION_EXCL_VISIT
) %>% glm(
  fy19_first_visit ~
    PR_SCORE +
    ID_SCORE +
    log10(EVALUATION_LOWER_BOUND + 1) +
    sqrt(ALLPFY_QUAL_EXCL_VISIT) +
    sqrt(PFY1_OUTREACH + PFY2_OUTREACH + PFY3_OUTREACH) +
    sqrt(PFY1_EMAIL + PFY2_EMAIL + PFY3_EMAIL) +
    sqrt(PFY1_PHONE + PFY2_PHONE + PFY3_PHONE) +
    PROGRAM_GROUP
  , data = .
)

summary(model_b)
```

Compare this to just using PR score.

```{r}
update(model_b, formula = . ~ . - ID_SCORE, data = visits_first_potential %>% mutate(
  ALLPFY_QUAL_EXCL_VISIT = PFY1_QUALIFICATION_EXCL_VISIT + PFY2_QUALIFICATION_EXCL_VISIT + PFY3_QUALIFICATION_EXCL_VISIT
)) %>% summary()
```

Interesting; ID_SCORE is far and away the best predictor of visits. Presumably PR score turns out to be too similar. let's look at visit proportion by ID score instead of PR score.

```{r}
visits_first_potential %>%
  # Only people with qualification attempts in the last couple years
  filter(qual_EXCL_VISIT | outreach) %>%
  ggplot(aes(x = ID_SCORE)) +
  geom_point(aes(y = as.numeric(fy19_first_visit), color = fy19_first_visit)) +
  geom_density(color = 'darkgray') +
  geom_smooth(aes(y = as.numeric(fy19_first_visit)), method = 'loess') +
  labs(y = 'First visit proportion', title = 'FY18-19 qualification attempts leading to FY19 first visit (ID score)')
```

Or looking at only the ID_score below 4.5:

```{r}
visits_first_potential %>%
  # Remove ID_SCORE above 4.5
  filter(ID_SCORE < 4.5) %>%
  # Only people with qualification attempts in the last couple years
  filter(qual_EXCL_VISIT | outreach) %>%
  ggplot(aes(x = ID_SCORE)) +
  geom_point(aes(color = fy19_first_visit, y = as.numeric(fy19_first_visit))) +
  geom_density(color = 'darkgray') +
  geom_smooth(aes(y = as.numeric(fy19_first_visit)), method = 'loess') +
  labs(y = 'First visit proportion', title = 'FY18-19 qualification attempts leading to FY19 first visit (ID score)')
```

That's a pretty compelling demonstration. Past 2.5 a higher ID_SCORE is pretty much linearly associated with a higher proportion of successful visits.

# PR and ID score metrics

Some summary statistics for each score.

```{r}
visits_first_potential %>%
  filter(qual_EXCL_VISIT | outreach) %>%
  select(PR_SCORE, ID_SCORE) %>%
  summary()
```

```{r}
binned_scores <- visits_first_potential %>% mutate(
  # Truncate in .5 increments
  PR_bin = trunc(PR_SCORE / .5) * .5
    , ID_bin = trunc(ID_SCORE / .5) * .5
)
```


Looking at the same population of potential first-time visits:

```{r}
binned_scores %>%
  filter(qual_EXCL_VISIT | outreach) %>%
  ggplot(aes(x = ID_SCORE, y = PR_SCORE)) +
  geom_point(aes(color = fy19_first_visit)) +
  geom_smooth(method = 'loess') +
  geom_smooth(method = 'lm') +
  geom_abline(slope = 1, intercept = 0, color = 'darkgray')
```

```{r}
binned_scores %>%
  filter(qual_EXCL_VISIT | outreach) %>%
  ggplot(aes(x = ID_SCORE, y = PR_SCORE)) +
  geom_point(aes(color = fy19_first_visit)) +
  geom_smooth(method = 'loess') +
  geom_smooth(method = 'lm') +
  geom_abline(slope = 1, intercept = 0, color = 'darkgray') +
  facet_grid(. ~ fy19_first_visit)
```

Binned:

```{r}
binned_scores %>%
  filter(qual_EXCL_VISIT | outreach) %>%
  group_by(PR_bin, ID_bin) %>%
  summarise(n = length(ID_NUMBER)) %>%
  ggplot(aes(x = ID_bin, y = PR_bin, fill = n)) +
  geom_tile(alpha = .75) +
  geom_text(aes(label = n), color = 'white', size = 3) +
  coord_equal() +
  xlim(c(-.5, 5)) +
  ylim(c(-.5, 5)) +
  scale_fill_gradient(name = 'count', trans = 'log10')
```

My basic takeaway is that the ID score is by far the best predictor of whether someone will take a visit, doing a much better job than any other variable in isolation (the "Excel sort" standard practice). ID scores above 2.5 or so start looking promising. However, we still need to think about whether taking a visit or expected gift value is the right metric.
