---
title: "CATracks data format"
output:
  html_document:
    df_print: paged
    toc: yes
  html_notebook:
    toc: yes
    toc_float:
      collapsed: no
---

# Goal

Format an exported KSM MG modeled data file into one that can be mass uploaded into the segments table.

# Setup

```{r setup, message = FALSE, warning = FALSE}
# Libraries
library(tidyverse)
```

```{r}
# Import data
predictions <- read.csv(file = 'data/2019-10-17 export.csv', stringsAsFactors = FALSE) %>%
  mutate(
    ID_NUMBER = str_pad(ID_NUMBER, width = 10, side = 'left', pad = '0')
    , n = 1
  )
```

# Thresholding

```{r}
predictions %>%
  group_by(est.giving.bin.coef1) %>%
  summarise(
    total = sum(n)
  ) %>%
  arrange(desc(est.giving.bin.coef1)) %>%
  mutate(
    cumsum = cumsum(total)
  )
```

`est.giving.bin` becomes the Kellogg Major Gift Identification Model (KMID1, KMID2, KMID3). 4+ = high, 3 = middle, 2- = low.

```{r}
predictions %>%
  group_by(ev.bin.coef1) %>%
  summarise(
    total = sum(n)
  ) %>%
  arrange(desc(ev.bin.coef1)) %>%
  mutate(
    cumsum = cumsum(total)
  )
```

`ev.bin` becomes the Kellogg Major Gift Prioritization Model (KMPR1, KMPR2, KMPR3). 3+ = high, 1-2 = medium, 0 = low.

# Segments

```{r}
output <- predictions %>% mutate(
  model1 = 'ID'
  , model2 = 'PR'
  , IDseg = case_when(
    est.giving.bin.coef1 >= 4 ~ 'KMID1'
    , est.giving.bin.coef1 == 3 ~ 'KMID2'
    , est.giving.bin.coef1 <= 2 ~ 'KMID3'
  )
  , PRseg = case_when(
    ev.bin.coef1 >= 3 ~ 'KMPR1'
    , ev.bin.coef1 %in% 1:2 ~ 'KMPR2'
    , ev.bin.coef1 == 0 ~ 'KMPR3'
  )
  , SEGMENT_YEAR = '2019'
  , SEGMENT_MONTH = '10'
  , IDxc = est.giving.coef1
  , PRxc = ev.coef1
  , DATA_SOURCE_CODE = 'KSM'
)

# write.csv(output, file = 'test.csv', row.names = FALSE)
```

# Segments to rows

```{r}
transformed <- output %>%
  gather('model', 'name', model1:model2) %>%
  mutate(
    SEGMENT_CODE = case_when(
      name == 'ID' ~ IDseg
      , name == 'PR' ~ PRseg
    )
    , XCOMMENT = case_when(
      name == 'ID' ~ IDxc
      , name == 'PR' ~ PRxc
    )
  ) %>% select(
    est.giving.coef1
    , est.p.coef1
    , ev.coef1
    , est.giving.bin.coef1
    , ev.bin.coef1
    , n
    , ID_NUMBER
    , SEGMENT_CODE
    , SEGMENT_YEAR
    , SEGMENT_MONTH
    , XCOMMENT
    , DATA_SOURCE_CODE
  ) %>% na.omit()
```

Verify results.

```{r}
transformed %>%
  filter(SEGMENT_CODE %in% c('KMPR1', 'KMPR2', 'KMPR3')) %>%
  group_by(ev.bin.coef1, SEGMENT_CODE) %>%
  summarise(
    total = sum(n)
  ) %>%
  arrange(desc(ev.bin.coef1))
```

```{r}
transformed %>%
  filter(SEGMENT_CODE %in% c('KMID1', 'KMID2', 'KMID3')) %>%
  group_by(est.giving.bin.coef1, SEGMENT_CODE) %>%
  summarise(
    total = sum(n)
  ) %>%
  arrange(desc(est.giving.bin.coef1))
```

# Export

```{r}
transformed %>%
  select(-est.giving.coef1, -est.p.coef1, -ev.coef1, -est.giving.bin.coef1, -ev.bin.coef1, -n) %>%
  write.csv(file = 'data/2019-10-18 KSM segments.csv', row.names = FALSE)
```

