---
title: "00 featuretoolsR demo"
output: html_notebook
---

# Goal

Understand how featuretoolsR, the R implementation of Featuretools [Python documentation](https://docs.featuretools.com/) can be used with relational data to reproducibly create derived features. I expect the package can create more features and do it with greater efficiency compared to subquerying/aggregating in SQL.

```{r setup, message = FALSE, warning = FALSE}
library(tidyverse)
library(readxl)
library(featuretoolsR)
```

# Data

Raw alumni and committee data for testing.

```{r}
degree <- read_xlsx(
  path = 'data/2020-02-21 entity committee test data.xlsx'
  , sheet = 1
  , guess_max = 1E6
)
committee <- read_xlsx(
  path = 'data/2020-02-21 entity committee test data.xlsx'
  , sheet = 2
  , guess_max = 1E6
)
```

Some data cleanup.

```{r}
committee <- committee %>%
  mutate(
    uid = paste(ID_NUMBER, XSEQUENCE) # unique identifier
  )
```


Create an entity set, i.e. list of related tables.

```{r}
es <- as_entityset(
  degree # data frame
  , entity_id = 'degree' # name of set
  , index = 'ID_NUMBER' # unique identifier
  , id = 'CATracks' # name of entity set
) %>% add_entity(
  committee
  , entity_id = 'committee'
  , index = 'uid'
  , time_index = 'START_DT_CALC' # when the record was created
)
```

Define relationships between the entities.

```{r}
es <- es %>% add_relationship(
  parent_set = 'degree'
  , child_set = 'committee'
  , parent_idx = 'ID_NUMBER'
  , child_idx = 'ID_NUMBER'
)
```

Examine the results.

```{r}
print(es)
```

# Primitives

# Deep feature synthesis