---
title: "02 Data file generation"
output:
  html_notebook:
    toc: TRUE
    toc_float:
      collapsed: FALSE
---

# Goal

Use featuretoolsR to create a large data frame for use in modeling.

```{r setup, message = FALSE, warning = FALSE}
library(tidyverse)
library(lubridate)
library(wranglR)
library(readxl)
library(foreach)
library(reticulate)
library(featuretoolsR)

source(file = 'code/featuretools helper functions.R')
```

# Data import

```{r}
fileloader <- function(filename, sheetnum) {
  read_xlsx(
    path = paste0('data/', filename)
    , sheet = sheetnum
    , guess_max = 1E7
  )
}

# Base table
households <- fileloader('2020-03-16 Engagement data.xlsx', 1)
# Biodata tables
bio_phone <- fileloader('2020-03-16 Addr phone.xlsx', 1)
bio_address <- fileloader('2020-03-16 Addr phone.xlsx', 2)
bio_employment <- fileloader('2020-03-16 employment email.xlsx', 1)
bio_email <- fileloader('2020-03-16 employment email.xlsx', 2)
# Engagement tables
eng_committee <- fileloader('2020-03-16 Engagement data.xlsx', 2)
eng_event <- fileloader('2020-03-16 Engagement data.xlsx', 3)
eng_activity <- fileloader('2020-03-16 Engagement data.xlsx', 4)
# Giving tables
giv_transactions <- fileloader('2020-03-16 Gift transactions.xlsx', 1)
giv_payments <- fileloader('2020-03-16 Gift data.xlsx', 1)
giv_giftclub <- fileloader('2020-03-16 Gift data.xlsx', 2)
# Prospect tables
prs_evaluation <- fileloader('2020-03-16 Prospect data.xlsx', 1)
prs_contact <- fileloader('2020-03-16 Prospect data.xlsx', 2)
prs_program <- fileloader('2020-03-16 Prospect data.xlsx', 3)
prs_assignment <- fileloader('2020-03-16 Prospect data.xlsx', 4)
```

# Data cleanup

```{r}
households <- households %>%
  # Factors
  mutate(
    RECORD_STATUS_CODE = factor(RECORD_STATUS_CODE)
    , HOUSEHOLD_PROGRAM = factor(HOUSEHOLD_PROGRAM)
    , HOUSEHOLD_PROGRAM_GROUP = factor(HOUSEHOLD_PROGRAM_GROUP)
  ) %>%
  # Numeric
  mutate_at(
    vars(ends_with('_YEAR'))
    , list(as.numeric)
  )

summary(households)
```

```{r}
bio_address <- bio_address %>%
  mutate(
    ADDR_TYPE_CODE = factor(ADDR_TYPE_CODE)
    , ADDR_TYPE = factor(ADDR_TYPE)
    , BUSINESS_IND = BUSINESS_IND == 'Y'
    , HBS_CODE = factor(HBS_CODE)
    , ADDR_STATUS_CODE = factor(ADDR_STATUS_CODE)
  )

summary(bio_address)
```

Note that any start year past 2020 is suspect.

```{r}
bio_email <- bio_email %>%
  mutate(
    EMAIL_TYPE_CODE = factor(EMAIL_TYPE_CODE)
    , EMAIL_TYPE = factor(EMAIL_TYPE)
    , EMAIL_STATUS_CODE = factor(EMAIL_STATUS_CODE)
  )

summary(bio_email)
```

```{r}
bio_employment <- bio_employment %>%
  mutate(
    JOB_STATUS_CODE = factor(JOB_STATUS_CODE)
    , SELF_EMPLOY_IND = SELF_EMPLOY_IND == 'Y'
    , MATCHING_STATUS_IND = MATCHING_STATUS_IND == 'Y'
  )

summary(bio_employment)
```

```{r}
bio_phone <- bio_phone %>%
  mutate(
    TELEPHONE_TYPE_CODE = factor(TELEPHONE_TYPE_CODE)
    , TELEPHONE_TYPE = factor(TELEPHONE_TYPE)
    , BUSINESS_IND = BUSINESS_IND == 'Y'
    , CURRENT_IND = CURRENT_IND == 'Y'
    , TELEPHONE_STATUS_CODE = factor(TELEPHONE_STATUS_CODE)
  )

summary(bio_phone)
```

Note to self - re-pull telephone data with additional types (done). Those 1902 dates do look pretty fishy.

```{r}
eng_activity <- eng_activity %>%
  mutate(
    ACTIVITY_DESC = factor(ACTIVITY_DESC)
    , KSM_ACTIVITY = !is.na(KSM_ACTIVITY)
    , ACTIVITY_PARTICIPATION_CODE = factor(ACTIVITY_PARTICIPATION_CODE)
  )

summary(eng_activity)
```

Clean up start/stop FY.

```{r}
eng_committee <- eng_committee %>%
  mutate(
    COMMITTEE_CODE = factor(COMMITTEE_CODE)
    , COMMITTEE_DESC = factor(COMMITTEE_DESC)
    , COMMITTEE_ROLE_CODE = factor(COMMITTEE_ROLE_CODE)
    , COMMITTEE_ROLE = factor(COMMITTEE_ROLE)
    , KSM_COMMITTEE = !is.na(KSM_COMMITTEE)
    , COMMITTEE_STATUS = factor(COMMITTEE_STATUS)
  )

summary(eng_committee)
```

```{r}
eng_event <- eng_event %>%
  mutate(
    EVENT_NAME = factor(EVENT_NAME)
    , KSM_EVENT = !is.na(KSM_EVENT)
    , EVENT_TYPE = factor(EVENT_TYPE)
  )

summary(eng_event)
```

```{r}
giv_giftclub <- giv_giftclub %>%
  mutate(
    GIFT_CLUB = factor(GIFT_CLUB)
    , GIFT_CLUB_CODE = factor(GIFT_CLUB_CODE)
    , GC_CATEGORY = factor(GC_CATEGORY)
  )

summary(giv_giftclub)
```

```{r}
giv_payments <- giv_payments %>%
  mutate(
    PLEDGE_FY = as.numeric(PLEDGE_FY)
  )

summary(giv_payments)
```

```{r}
giv_transactions <- giv_transactions %>%
  mutate(
    TRANSACTION_TYPE = factor(TRANSACTION_TYPE)
    , TX_GYPM_IND = factor(TX_GYPM_IND)
    , ASSOCIATED_DESC = factor(ASSOCIATED_DESC)
    , PAYMENT_TYPE = factor(PAYMENT_TYPE)
    , ALLOC_SHORT_NAME = factor(ALLOC_SHORT_NAME)
    , AF_FLAG = !is.na(AF_FLAG)
    , CRU_FLAG = !is.na(CRU_FLAG)
    , PROPOSAL_ID = factor(PROPOSAL_ID)
  )

summary(giv_transactions)
```

```{r}
prs_assignment <- prs_assignment %>%
  mutate(
    ASSIGNMENT_TYPE = factor(ASSIGNMENT_TYPE)
    , ASSIGNMENT_TYPE_DESC = factor(ASSIGNMENT_TYPE_DESC)
    , ASSIGNMENT_REPORT_NAME = factor(ASSIGNMENT_REPORT_NAME)
    , COMMITTEE_DESC = factor(COMMITTEE_DESC)
  )

prs_assignment %>% select(-ASSIGNMENT_REPORT_NAME) %>% summary()
```

No meaningful COMMITTEE_DESC data.

```{r}
prs_contact <- prs_contact %>%
  mutate(
    CONTACT_CREDIT_TYPE = factor(CONTACT_CREDIT_TYPE)
    , CONTACT_CREDIT_DESC = factor(CONTACT_CREDIT_DESC)
    , ARD_STAFF = !is.na(ARD_STAFF)
    , FRONTLINE_KSM_STAFF = !is.na(FRONTLINE_KSM_STAFF)
    , CONTACT_TYPE_CATEGORY = factor(CONTACT_TYPE_CATEGORY)
    , VISIT_TYPE = factor(VISIT_TYPE)
  )

summary(prs_contact)
```

```{r}
prs_evaluation <- prs_evaluation %>%
  mutate(
    EVALUATION_TYPE = factor(EVALUATION_TYPE)
    , EVALUATION_DESC = factor(EVALUATION_DESC)
    , RATING_CODE = factor(RATING_CODE)
    , RATING_DESC = factor(RATING_DESC)
  )

summary(prs_evaluation)
```

```{r}
prs_program <- prs_program %>%
  mutate(
    PROGRAM_CODE = factor(PROGRAM_CODE)
    , PROGRAM = factor(PROGRAM)
  )

summary(prs_program)
```

# Quick test

Test features with biodata only.

```{r}
hh <- households %>%
  select(
    -HOUSEHOLD_RPT_NAME
    , -HOUSEHOLD_SPOUSE_ID
    , -HOUSEHOLD_SPOUSE_RPT_NAME
    , -RECORD_STATUS_CODE
    , -DATA_AS_OF
  )
```

Create entities.

```{r}
es <- as_entityset(
  hh # data frame
  , entity_id = 'household' # name of entity
  , index = 'HOUSEHOLD_ID' # unique identifier
  , id = 'household' # name of entity set
) %>%
  # Bio entities
  add_entity(
    bio_address %>% select(-START_DT, -STOP_DT, -DATE_MODIFIED, -ADDR_TYPE_CODE)
    , entity_id = 'bio_address'
    , make_index = TRUE
    , index = 'bio_address_idx'
    , time_index = 'DATE_ADDED' # when the record can first be used
  ) %>%
  add_entity(
  bio_email %>% select(-EMAIL_TYPE_CODE, -START_DT, -STOP_DT)
  , entity_id = 'bio_email'
  , make_index = TRUE
  , index = 'bio_email_idx'
  , time_index = 'DATE_ADDED' # when the record can first be used
  ) %>%
  add_entity(
  bio_employment %>% select(-START_DT, -STOP_DT, -JOB_TITLE, -EMPLOYER_NAME)
  , entity_id = 'bio_employment'
  , make_index = TRUE
  , index = 'bio_employment_idx'
  , time_index = 'DATE_ADDED' # when the record can first be used
  ) %>%
  add_entity(
  bio_phone %>% select(-TELEPHONE_TYPE_CODE, -START_DT, -STOP_DT)
  , entity_id = 'bio_phone'
  , make_index = TRUE
  , index = 'bio_phone_idx'
  , time_index = 'DATE_ADDED' # when the record can first be used
)
```

Create relationships.

```{r}
es <- es %>%
  add_relationship(
    parent_set = 'household'
    , child_set = 'bio_address'
    , parent_idx = 'HOUSEHOLD_ID'
    , child_idx = 'HOUSEHOLD_ID'
  ) %>%
  add_relationship(
    parent_set = 'household'
    , child_set = 'bio_email'
    , parent_idx = 'HOUSEHOLD_ID'
    , child_idx = 'HOUSEHOLD_ID'
  ) %>%
  add_relationship(
    parent_set = 'household'
    , child_set = 'bio_employment'
    , parent_idx = 'HOUSEHOLD_ID'
    , child_idx = 'HOUSEHOLD_ID'
  ) %>%
  add_relationship(
    parent_set = 'household'
    , child_set = 'bio_phone'
    , parent_idx = 'HOUSEHOLD_ID'
    , child_idx = 'HOUSEHOLD_ID'
  )
```

```{r}
print(es)
```

```{r}
# Use base R data
dfs_r <- es %>% dfs(
  target_entity = 'household'
  , agg_primitives = c('count')
  , trans_primitives = c('month', 'year')
  , max_depth = 2
  , verbose = TRUE
)
```

```{r, rows.print = 20}
print_features(dfs_r)
```

We see that dfs created counts for the total number of address, email, etc. entries on the table per household_id. However, it would be more useful to get counts broken down by other criteria -- for example, how many historical home addresses, active emails, etc.

# "Interesting values" function

The featuretools "interesting values" functionality creates new aggregation fields for each of a predefined set of factor levels -- or just the n most common.

```{python}
import featuretools as ft

# Import R data
py_es = r.es

# Add interesting values tabulation; here I'm looking at up to the 6 most common levels for each table.
py_es.add_interesting_values(max_values = 6, verbose = True)
```

```{r}
# Import data from python
dfs_py <- py$py_es %>% dfs(
  target_entity = 'household'
  , agg_primitives = c('count')
  , trans_primitives = c('month', 'year')
  , max_depth = 2
  , verbose = TRUE
)
```

```{r, rows.print = 20}
print_features(dfs_py)
```

Here we see the interesting values functionality added counts specifically by contact status, type, etc. This took only a few seconds, which definitely beats doing this for dozens of fields by hand.

# Full run

## Load data

```{r, warning = FALSE}
# Change this to the date the data files were generated
data_dt <- '2020-04-08'
source(file = 'code/data xlsx import.R')
```

## Check data

```{r}
source(file = 'code/data validation.R')
```

## Build entity set

```{r}
# r_entityset <- as_entityset(
#   hh # data frame
#   , entity_id = 'household' # name of entity
#   , index = 'HOUSEHOLD_ID' # unique identifier
#   , time_index = 'DATE_ADDED' # when the record can first be used
#   , id = 'household' # name of entity set
# ) %>%
# # Bio entities
#   add_entity(
#     bio_address %>% select(-START_DT, -STOP_DT, -DATE_MODIFIED, -ADDR_TYPE_CODE)
#     , entity_id = 'bio_address'
#     , make_index = TRUE
#     , index = 'bio_address_idx'
#     , time_index = 'DATE_ADDED' # when the record can first be used
#   ) %>%
#   add_entity(
#   bio_email %>% select(-EMAIL_TYPE_CODE, -START_DT, -STOP_DT)
#   , entity_id = 'bio_email'
#   , make_index = TRUE
#   , index = 'bio_email_idx'
#   , time_index = 'DATE_ADDED' # when the record can first be used
#   ) %>%
#   add_entity(
#   bio_employment %>% select(-START_DT, -STOP_DT, -JOB_TITLE, -EMPLOYER_NAME)
#   , entity_id = 'bio_employment'
#   , make_index = TRUE
#   , index = 'bio_employment_idx'
#   , time_index = 'DATE_ADDED' # when the record can first be used
#   ) %>%
#   add_entity(
#   bio_phone %>% select(-TELEPHONE_TYPE_CODE, -START_DT, -STOP_DT)
#   , entity_id = 'bio_phone'
#   , make_index = TRUE
#   , index = 'bio_phone_idx'
#   , time_index = 'DATE_ADDED' # when the record can first be used
#   ) %>%
# # Engagement entities
#   add_entity(
#   eng_activity %>% select(-ID_NUMBER, -START_DT, -STOP_DT)
#   , entity_id = 'eng_activity'
#   , make_index = TRUE
#   , index = 'eng_activity_idx'
#   , time_index = 'DATE_ADDED'
#   ) %>%
#   add_entity(
#   eng_committee %>% select(-ID_NUMBER, -COMMITTEE_CODE, -COMMITTEE_ROLE_CODE, -COMMITTEE_ROLE_XSEQUENCE)
#   , entity_id = 'eng_committee'
#   , make_index = TRUE
#   , index = 'eng_committee_idx'
#   , time_index = 'START_DT_CALC'
#   ) %>%
#   add_entity(
#   eng_event %>% select(-ID_NUMBER, -EVENT_ID)
#   , entity_id = 'eng_event'
#   , make_index = TRUE
#   , index = 'eng_event_idx'
#   , time_index = 'START_DT'
#   ) %>%
# # Giving entities
#   add_entity(
#   giv_giftclub %>% select(-GIFT_CLUB_CODE, -GIFT_CLUB_START_DATE, -GIFT_CLUB_END_DATE, -STOP_DT)
#   , entity_id = 'giv_giftclub'
#   , make_index = TRUE
#   , index = 'giv_giftclub_idx'
#   , time_index = 'START_DT'
#   ) %>%
#   add_entity(
#   giv_transactions %>% select(-TX_NUMBER, -TX_SEQUENCE, -MATCHED_TX_NUMBER, -PLEDGE_NUMBER)
#   , entity_id = 'giv_transactions'
#   , make_index = TRUE
#   , index = 'giv_transactions_idx'
#   , time_index = 'DATE_OF_RECORD'
#   )
# # Prospect entities
#   add_entity(
#   prs_assignment %>% select(-ASSIGNMENT_TYPE, COMMITTEE_DESC)
#   , entity_id = 'prs_assignment'
#   , make_index = TRUE
#   , index = 'prs_assignment_idx'
#   , time_index = 'START_DT_CALC'
#   ) %>%
#   add_entity(
#   prs_contact %>% select(-CONTACT_CREDIT_TYPE)
#   , entity_id = 'prs_contact'
#   , make_index = TRUE
#   , index = 'prs_contact_idx'
#   , time_index = 'CONTACT_DATE'
#   ) %>%
#   add_entity(
#   prs_evaluation %>% select(-EVALUATION_TYPE, -RATING_CODE)
#   , entity_id = 'prs_evaluation'
#   , make_index = TRUE
#   , index = 'prs_evaluation_idx'
#   , time_index = 'EVAL_DT'
#   ) %>%
#   add_entity(
#   prs_program %>% select(-PROGRAM_CODE, -START_DATE, -STOP_DATE)
#   , entity_id = 'prs_program'
#   , make_index = TRUE
#   , index = 'prs_program_idx'
#   , time_index = 'DATE_ADDED'
#   )
```

Create relationships.

```{r}
# r_entityset <- r_entityset %>%
# # Bio
#   add_relationship(
#     parent_set = 'household'
#     , child_set = 'bio_address'
#     , parent_idx = 'HOUSEHOLD_ID'
#     , child_idx = 'HOUSEHOLD_ID'
#   ) %>%
#   add_relationship(
#     parent_set = 'household'
#     , child_set = 'bio_email'
#     , parent_idx = 'HOUSEHOLD_ID'
#     , child_idx = 'HOUSEHOLD_ID'
#   ) %>%
#   add_relationship(
#     parent_set = 'household'
#     , child_set = 'bio_employment'
#     , parent_idx = 'HOUSEHOLD_ID'
#     , child_idx = 'HOUSEHOLD_ID'
#   ) %>%
#   add_relationship(
#     parent_set = 'household'
#     , child_set = 'bio_phone'
#     , parent_idx = 'HOUSEHOLD_ID'
#     , child_idx = 'HOUSEHOLD_ID'
#   ) %>%
# # Engagement
#   add_relationship(
#     parent_set = 'household'
#     , child_set = 'eng_activity'
#     , parent_idx = 'HOUSEHOLD_ID'
#     , child_idx = 'HOUSEHOLD_ID'
#   ) %>%
#   add_relationship(
#     parent_set = 'household'
#     , child_set = 'eng_committee'
#     , parent_idx = 'HOUSEHOLD_ID'
#     , child_idx = 'HOUSEHOLD_ID'
#   ) %>%
#   add_relationship(
#     parent_set = 'household'
#     , child_set = 'eng_event'
#     , parent_idx = 'HOUSEHOLD_ID'
#     , child_idx = 'HOUSEHOLD_ID'
#   ) %>%
# # Giving
#   add_relationship(
#     parent_set = 'household'
#     , child_set = 'giv_giftclub'
#     , parent_idx = 'HOUSEHOLD_ID'
#     , child_idx = 'HOUSEHOLD_ID'
#   ) %>%
#   add_relationship(
#     parent_set = 'household'
#     , child_set = 'giv_transactions'
#     , parent_idx = 'HOUSEHOLD_ID'
#     , child_idx = 'HOUSEHOLD_ID'
#   ) %>%
# # Prospect
```

```{r}
# print(r_entityset)
```

## Compute features

```{r}
# features <- r_entityset %>% dfs(
#    target_entity = 'household'
#    , agg_primitives = c('count', 'sum', 'std', 'mean', 'max', 'min', 'median', 'last', 'percent_true')
#    , trans_primitives = c('month', 'year', 'subtract_numeric', 'divide_numeric', 'time_since_previous')
#    , max_depth = 2
#    , verbose = TRUE
# )
```

```{r, rows.print = 25}
# print_features(features)
```

