---
title: "Annual Fund $10K Model Data Exploration"
output: html_notebook
---

# Goal

Clean up and explore the dataset generated with the `ksm-af-10k-data.sql` query with reusable R code.

# Setup

## R version

```{r get.sessionInfo}
sessionInfo()
```

## Required packages

My code makes use of several packages:

```{r get.packages}
scan('PACKAGES.txt', what = 'character')
```

This code will install and load them as necessary. First, install `devtools` and my `wranglR` package:

```{r get.wranglR}
# Check to see if devtools is unavailable
if (!('devtools' %in% utils::installed.packages()[, 'Package'])) {
  # If unavailable, install it
  install.packages('devtools')
}

# Check to see if wranglR is unavailable
if (!('wranglR' %in% utils::installed.packages()[, 'Package'])) {
  # If unavailable, use devtools to install it via GitHub
  devtools::install_github('phively/wranglR')
}
```

Now load `wranglR` and use it to install or load the other required packages:

```{r load.packages, warning=FALSE, message=FALSE}
# Load wranglR
library(wranglR)

# Feed the libraries specified in PACKAGES.txt into wranglR's Libraries() function
Libraries(scan(file = 'PACKAGES.txt', what = 'character'))
```

## Data preparation

```{r load.data}
### Import the data
full.data <- read.csv('data/2017-11-29 AF 10K Model.csv', stringsAsFactors = FALSE, strip.white = TRUE,
                      colClasses = c('ID_NUMBER' = 'character',
                                     'HOUSEHOLD_ID' = 'character',
                                     'BIRTH_DT' = 'character')) %>%
  # Drop any null rows
  filter(!is.na(RN)) %>%
  # Drop row numbers
  select(-RN)

### Clean the data
full.data <- full.data %>%
  # Convert NA to 0
  mutate_all(
    funs(
      ifelse(is.na(.), 0, .)
    )
  ) %>%
  # Convert to date
  mutate(
      DOB = ToDate(BIRTH_DT, method = 'ymd')
  ) %>%
  # Convert certain character to factor
  mutate(
    # Factors
      RECORD_STATUS_CODE = factor(RECORD_STATUS_CODE) %>% relevel(ref = 'A') # Active
    , HOUSEHOLD_RECORD = factor(HOUSEHOLD_RECORD) %>% relevel(ref = 'AL') # Alumni
    , PROGRAM_GROUP = factor(PROGRAM_GROUP) %>% relevel(ref = 'FT') # Full-Time
    , PREF_ADDR_TYPE_CODE = factor(PREF_ADDR_TYPE_CODE) %>% relevel(ref = 'H') # Home
    , HOUSEHOLD_STATE = factor(HOUSEHOLD_STATE) %>% relevel(ref = 'IL') # Illinois
    , HOUSEHOLD_COUNTRY = factor(HOUSEHOLD_COUNTRY) %>% relevel(ref = 'United States')
    , HOUSEHOLD_CONTINENT = factor(HOUSEHOLD_CONTINENT) %>% relevel(ref = 'North America')
    , EVALUATION_RATING = factor(EVALUATION_RATING) %>% relevel(ref = '') # Unrated
    # Indicators
    , BUS_IS_EMPLOYED = factor(BUS_IS_EMPLOYED == 'Y')
    , BUS_HIGH_LVL_JOB_TITLE = factor(BUS_HIGH_LVL_JOB_TITLE != '')
    , BUS_CAREER_SPEC_FINANCE = factor(BUS_CAREER_SPEC_FINANCE != '')
    , BUS_GIFT_MATCH = factor(BUS_GIFT_MATCH == 'Y')
    , HAS_HOME_ADDR = factor(HAS_HOME_ADDR == 'Y')
    , HAS_ALT_HOME_ADDR = factor(HAS_ALT_HOME_ADDR == 'Y')
    , HAS_BUS_ADDR = factor(HAS_BUS_ADDR == 'Y')
    , HAS_SEASONAL_ADDR = factor(HAS_SEASONAL_ADDR == 'Y')
    , HAS_HOME_PHONE = factor(HAS_HOME_PHONE == 'Y')
    , HAS_BUS_PHONE = factor(HAS_BUS_PHONE == 'Y')
    , HAS_MOBILE_PHONE = factor(HAS_MOBILE_PHONE == 'Y')
    , HAS_HOME_EMAIL = factor(HAS_HOME_EMAIL == 'Y')
    , HAS_BUS_EMAIL = factor(HAS_BUS_EMAIL == 'Y')
    , KSM_PROSPECT_ACTIVE = factor(KSM_PROSPECT_ACTIVE == 'Y')
    , KSM_PROSPECT_ANY = factor(KSM_PROSPECT_ANY == 'Y')
  )
```

"Failed to parse" means `DOB` is NA due to invalid `BIRTH_DT` format.

# Data Exploration

