---
title: "Annual Fund $10K Model Data Exploration"
output: html_notebook
---

# Goal

Clean up and explore the dataset generated with the `ksm-af-10k-data.sql` query with reusable R code.

# Setup

## R version

```{r get.sessionInfo}
sessionInfo()
```

## Required packages

My code makes use of several packages:

```{r get.packages}
scan('PACKAGES.txt', what = 'character')
```

This code will install and load them as necessary. First, install `devtools` and my `wranglR` package:

```{r get.wranglR}
# Check to see if devtools is unavailable
if (!('devtools' %in% utils::installed.packages()[, 'Package'])) {
  # If unavailable, install it
  install.packages('devtools')
}

# Check to see if wranglR is unavailable
if (!('wranglR' %in% utils::installed.packages()[, 'Package'])) {
  # If unavailable, use devtools to install it via GitHub
  devtools::install_github('phively/wranglR')
}
```

## Load libraries

Now load `wranglR` and use it to install or load the other required packages:

```{r load.packages, warning=FALSE, message=FALSE}
# Load wranglR
library(wranglR)

# Feed the libraries specified in PACKAGES.txt into wranglR's Libraries() function
Libraries(scan(file = 'PACKAGES.txt', what = 'character'))
```

## Data preparation

```{r load.data}
### Import the data
full.data <- read.csv('data/2017-12-21 AF 10K Model.csv', stringsAsFactors = FALSE, strip.white = TRUE,
                      colClasses = c('ID_NUMBER' = 'character'
                                     , 'HOUSEHOLD_ID' = 'character'
                                     , 'BIRTH_DT' = 'character'
                                     , 'ENTITY_DT_ADDED' = 'character'
                                     , 'GIVING_MAX_CASH_DT' = 'character'
                                     , 'GIVING_MAX_PLEDGE_DT' = 'character')) %>%
  # Drop any null rows
  filter(!is.na(RN)) %>%
  # Drop row numbers
  select(-RN)

### Clean the data
full.data <- full.data %>%
  # Convert NA to 0
  mutate_all(
    funs(
      ifelse(is.na(.), 0, .)
    )
  ) %>%
  # Convert to date
  mutate(
      DOB = ToDate(BIRTH_DT, method = 'ymd')
    , ENTITY_DT = ToDate(ENTITY_DT_ADDED, method = 'mdy')
    , GIVING_MAX_CASH_DT = ToDate(GIVING_MAX_CASH_DT, method = 'mdy')
    , GIVING_MAX_PLEDGE_DT = ToDate(GIVING_MAX_CASH_DT, method = 'mdy')
  ) %>%
  # Extract dates
  mutate(
      GIVING_MAX_CASH_YR = year(GIVING_MAX_CASH_DT)
    , GIVING_MAX_CASH_MO = month(GIVING_MAX_CASH_DT) %>% factor()
  ) %>%
  # Convert certain character to factor
  mutate(
    # Factors
      RECORD_STATUS_CODE = factor(RECORD_STATUS_CODE) %>% relevel(ref = 'A') # Active
    , HOUSEHOLD_RECORD = factor(HOUSEHOLD_RECORD) %>% relevel(ref = 'AL') # Alumni
    , PROGRAM_GROUP = factor(PROGRAM_GROUP) %>% relevel(ref = 'FT') # Full-Time
    , PREF_ADDR_TYPE_CODE = factor(PREF_ADDR_TYPE_CODE) %>% relevel(ref = 'H') # Home
    , HOUSEHOLD_STATE = factor(HOUSEHOLD_STATE) %>% relevel(ref = 'IL') # Illinois
    , HOUSEHOLD_COUNTRY = factor(HOUSEHOLD_COUNTRY) %>% relevel(ref = 'United States')
    , HOUSEHOLD_CONTINENT = factor(HOUSEHOLD_CONTINENT) %>% relevel(ref = 'North America')
    , EVALUATION_RATING = factor(EVALUATION_RATING) %>% relevel(ref = '') # Unrated
    # Indicators
    , BUS_IS_EMPLOYED = factor(BUS_IS_EMPLOYED == 'Y')
    , BUS_HIGH_LVL_JOB_TITLE = factor(BUS_HIGH_LVL_JOB_TITLE != '')
    , BUS_CAREER_SPEC_FINANCE = factor(BUS_CAREER_SPEC_FINANCE != '')
    , BUS_GIFT_MATCH = factor(BUS_GIFT_MATCH == 'Y')
    , HAS_HOME_ADDR = factor(HAS_HOME_ADDR == 'Y')
    , HAS_ALT_HOME_ADDR = factor(HAS_ALT_HOME_ADDR == 'Y')
    , HAS_BUS_ADDR = factor(HAS_BUS_ADDR == 'Y')
    , HAS_SEASONAL_ADDR = factor(HAS_SEASONAL_ADDR == 'Y')
    , HAS_HOME_PHONE = factor(HAS_HOME_PHONE == 'Y')
    , HAS_BUS_PHONE = factor(HAS_BUS_PHONE == 'Y')
    , HAS_MOBILE_PHONE = factor(HAS_MOBILE_PHONE == 'Y')
    , HAS_HOME_EMAIL = factor(HAS_HOME_EMAIL == 'Y')
    , HAS_BUS_EMAIL = factor(HAS_BUS_EMAIL == 'Y')
    , KSM_PROSPECT_ACTIVE = factor(KSM_PROSPECT_ACTIVE == 'Y')
    , KSM_PROSPECT_ANY = factor(KSM_PROSPECT_ANY == 'Y')
  )
```

"Failed to parse" means `DOB` is NA due to invalid `BIRTH_DT` format.

# Data Exploration

## Diagnostics

**Observations** from `summary(full.data)`

  * Need to be cautious about the inserted 0s; e.g. discretize years?
  * Should gift amounts be inflation-indexed? (CPI-U adjustment)
  * DOB missing and impossible values: impute?

```{r plot.grad.year}
full.data %>% filter(FIRST_KSM_YEAR != 0) %>%
  ggplot(aes(x = FIRST_KSM_YEAR, fill = RECORD_STATUS_CODE)) + geom_histogram(alpha = .5, binwidth = 1) +
  facet_grid(RECORD_STATUS_CODE ~ .) + scale_y_log10()
```

  * Check 1800 grad year
  
```{r plot.dob}
full.data %>% filter(!is.na(DOB) & year(DOB) != 0) %>%
  ggplot(aes(x = year(DOB), fill = RECORD_STATUS_CODE)) + geom_histogram(alpha = .5, binwidth = 1) +
  facet_grid(RECORD_STATUS_CODE ~ .) + scale_y_log10()
```

  * Check the person with DOB year 0
  * Check people with DOB after 2000...or in general after they graduated?

## Inflation adjustment exploration

Plot largest nominal cash gift from each donor against the year in which it was made.

```{r plot.amt.by.year}
full.data %>% filter(GIVING_MAX_CASH_AMT > 0) %>%
  ggplot(aes(x = GIVING_MAX_CASH_YR, y = GIVING_MAX_CASH_AMT)) +
  geom_point() + geom_jitter(width = .25, height = 0) +
  scale_y_log10(breaks = 10^(-2:7)) + geom_smooth() +
   # Red line is linear trend
  geom_smooth(method = 'lm', color = 'red') +
  labs(title = "Nominal dollars by year")
```

Slight upward slope over the last 30-40 years.

```{r infl.adjust}
# Load historical CPI-U data
cpi.u <- read.csv('data/CPI 2017-09-01.csv', strip.white = TRUE) %>%
  mutate(
    AS_OF = mdy(AS_OF)
    , YEAR = CALENDAR_YEAR
  )
cpi.max <- cpi.u %>% filter(CALENDAR_YEAR == max(CALENDAR_YEAR)) %>% select(CPI_U_AVG) %>% as.numeric()
```

Plot largest inflation-adjusted cash gift (Sept 2017 dollars) from each donor against the year in which it was made.

```{r plot.cpiu.amt.by.year}
full.data %>% filter(GIVING_MAX_CASH_AMT > 0) %>%
  # Join CPI-U columns
  mutate(YEAR = GIVING_MAX_CASH_YR) %>%
  inner_join(cpi.u, by = c('YEAR')) %>%
  # Compute inflation-adjusted amount
  mutate(CPIU_MAX_CASH_AMT = GIVING_MAX_CASH_AMT * cpi.max/CPI_U_AVG) %>%
  # Plot results
  ggplot(aes(x = YEAR, y = CPIU_MAX_CASH_AMT)) +
  geom_point() + geom_jitter(width = .25, height = 0) +
  scale_y_log10(breaks = 10^(-2:7)) + geom_smooth() +
  # Red line is linear trend
  geom_smooth(method = 'lm', color = 'red') +
  labs(title = "September 2017 dollars by year")
```

Inflation-adjusted there's still a slight upward slope over time. Does look remarkably consistent over time.

## Response variable transformations

For Annual Giving, largest single cash gift is the logical response variable.

```{r summary2}
summary2 <- function(x) {
  summary(x) %>% c(
    StDev = sd(x)
    , Skewnewss = e1071::skewness(x)
    , Kurtosis = e1071::kurtosis(x)
  ) %>% signif(digits = 3)
}
```
```{r rv.stats}
full.data$GIVING_MAX_CASH_AMT %>% summary2()
# Log10 transformation, no 0
{full.data %>% filter(GIVING_MAX_CASH_AMT > 0)}$GIVING_MAX_CASH_AMT %>% log10() %>% summary2()
# Log10 transformation, with 0
{full.data$GIVING_MAX_CASH_AMT + 1} %>% log10() %>% summary2()
```

Giving is left-truncated at 0. I'd previously looked into alternate distributions with different datasets but the additional complexity was not worth it.

```{r rv.trans.plot}
# Density histogram
full.data %>% filter(GIVING_MAX_CASH_AMT > 0) %>%
  ggplot(aes(x = GIVING_MAX_CASH_AMT)) +
  geom_histogram(aes(y = ..density..), bins = 50, alpha = .5) +
  geom_density() +
  scale_x_log10(breaks = 10^(-2:7),
                minor_breaks = c(.25 * 10^(-2:7), .5 * 10^(-2:7), .75 * 10^(-2:7)),
                labels = scales::dollar) +
  theme(panel.grid.minor.x = element_line(linetype = 'dashed')) +
  labs(title = 'No 0')
```

Log scale appears to be a decent choice, though it's still right-skewed.

```{r rv.boxcox}
# Box-Cox analysis
bc <- lm(log10(GIVING_MAX_CASH_AMT) ~ GIVING_MAX_CASH_YR, data = full.data %>% filter(GIVING_MAX_CASH_AMT > 1)) %>%
  MASS::boxcox()
bc$x[which(bc$y == max(bc$y))]
```

According to this the best transformation on the 1-variable model is approximately $\sqrt[4]{\log(x)}$, which is quite hard to interpret. Log transformation is likely sufficient and has a reasonable explanation (error term $\epsilon$ enters multiplicatively).

## Response variable correlations

### Functions

```{r rv.plotter}
# ggplot2 boxplotting function from colname
boxplotter <- function(cname, data = full.data) {
  data %>% select(GIVING_MAX_CASH_AMT) %>% cbind(x = data[, cname]) %>%
    filter(GIVING_MAX_CASH_AMT > 0) %>%
    ggplot(aes(x = x, y = GIVING_MAX_CASH_AMT)) +
    geom_boxplot(alpha = .6) +
    scale_y_log10(breaks = 10^(-2:7), labels = scales::dollar) +
    labs(x = cname)
}

# ggplot2 scatterplotting function from colname
scatterplotter <- function(cname, data = full.data) {
  data %>% select(GIVING_MAX_CASH_AMT) %>% cbind(x = data[, cname]) %>%
    filter(GIVING_MAX_CASH_AMT > 0) %>%
    ggplot(aes(x = x, y = GIVING_MAX_CASH_AMT)) +
    geom_point(alpha = .6) +
    geom_smooth() +
    geom_smooth(method = 'lm', color = 'red') +
    scale_y_log10(breaks = 10^(-2:7), labels = scales::dollar) +
    labs(x = cname)
}

# ggplot2 histogram function from colname
histogrammer <- function(cname, m = 30, data = full.data) {
  data %>% cbind(x = data[, cname]) %>%
    ggplot(aes(x = x)) +
    geom_histogram(alpha = .6, bins = m) +
    geom_density(aes(y = ..count..)) +
    labs(x = cname)
}
```


### Categorical variables

```{r rv.cat.rsc}
summary(full.data$RECORD_STATUS_CODE)
boxplotter('RECORD_STATUS_CODE')
```

Good number within each factor and a meaningful indicator.

```{r rv.cat.hhr}
summary(full.data$HOUSEHOLD_RECORD)
boxplotter('HOUSEHOLD_RECORD')
```

Only a few students; good number of nonalumni.
**Consider combining AL and ST**

```{r rv.cat.hhprim}
summary(full.data$HH_PRIMARY %>% as.factor())
boxplotter('HH_PRIMARY')
```

Should not include `N` in analysis as it duplicates a `Y` elsewhere in the file.
**EXCLUDE**

```{r rv.cat.pg}
summary(full.data$PROGRAM_GROUP)
boxplotter('PROGRAM_GROUP')
```

Only 20 UNK; others have a good number of rows.
**Consider dropping UNK or treating as nonalum**

```{r rv.cat.patc}
summary(full.data$PREF_ADDR_TYPE_CODE)
boxplotter('PREF_ADDR_TYPE_CODE')
```

Several with very few examples; a few options here. Could combine all the Home-like and Business-like, and leave an Other for seasonal etc. Could also roll all the < 500, say, into an Other.
**Consider combining categories**

```{r rv.cat.hhs}
summary(full.data$HOUSEHOLD_STATE)
boxplotter('HOUSEHOLD_STATE')
```

This appears to include Canadian and Australian states. Could also combine some of the lower-population areas.
**Consider combining, perhaps based on region**

```{r rv.cat.hhc}
summary(full.data$HOUSEHOLD_CONTINENT)
boxplotter('HOUSEHOLD_CONTINENT')
```

Clear difference between those with and without an address; probably fine to include in model to suss out other differences.

```{r rv.cat.bie}
summary(full.data$BUS_IS_EMPLOYED)
boxplotter('BUS_IS_EMPLOYED')
```

Looks fine.

```{r rv.cat.bhljt}
summary(full.data$BUS_HIGH_LVL_JOB_TITLE)
boxplotter('BUS_HIGH_LVL_JOB_TITLE')
```

Looks fine.

```{r rv.cat.bcsf}
summary(full.data$BUS_CAREER_SPEC_FINANCE)
boxplotter('BUS_CAREER_SPEC_FINANCE')
```

Looks fine.

```{r rv.cat.bgm}
summary(full.data$BUS_GIFT_MATCH)
boxplotter('BUS_GIFT_MATCH')
```

Very few that have matching companies; apparently this is not tracked consistently?
**EXCLUDE**

```{r rv.cat.hha}
summary(full.data$HAS_HOME_ADDR)
boxplotter('HAS_HOME_ADDR')
```

Looks fine.

```{r rv.cat.haha}
summary(full.data$HAS_ALT_HOME_ADDR)
boxplotter('HAS_ALT_HOME_ADDR')
```

Looks fine.

```{r rv.cat.hba}
summary(full.data$HAS_BUS_ADDR)
boxplotter('HAS_BUS_ADDR')
```

Looks fine.

```{r rv.cat.hsa}
summary(full.data$HAS_SEASONAL_ADDR)
boxplotter('HAS_SEASONAL_ADDR')
```

Low count.
**Consider combining with alternate home address**

```{r rv.cat.hhp}
summary(full.data$HAS_HOME_PHONE)
boxplotter('HAS_HOME_PHONE')
```

Looks fine.

```{r rv.cat.hbp}
summary(full.data$HAS_BUS_PHONE)
boxplotter('HAS_BUS_PHONE')
```

Looks fine.

```{r rv.cat.hmp}
summary(full.data$HAS_MOBILE_PHONE)
boxplotter('HAS_MOBILE_PHONE')
```

Looks fine.

```{r rv.cat.hhe}
summary(full.data$HAS_HOME_EMAIL)
boxplotter('HAS_HOME_EMAIL')
```

Looks fine.

```{r rv.cat.hbe}
summary(full.data$HAS_BUS_EMAIL)
boxplotter('HAS_BUS_EMAIL')
```

Looks fine.

```{r rv.cat.eval}
summary(full.data$EVALUATION_RATING)
boxplotter('EVALUATION_RATING')
```

Would likely combine all the $1M+ levels. Note that the evaluation rating process is causally associated with solicitations and thus gift size.
**CAUTION - endogenous**

```{r rv.cat.kpact}
summary(full.data$KSM_PROSPECT_ACTIVE)
boxplotter('KSM_PROSPECT_ACTIVE')
```

Looks fine.

```{r rv.cat.kpany}
summary(full.data$KSM_PROSPECT_ANY)
boxplotter('KSM_PROSPECT_ANY')
```

Looks fine.
**Consider making a new factor combining `KSM_PROSPECT_ACTIVE` and `KSM_PROSPECT_ANY`**

```{r rv.cat.kpanyall}
# Derived variable
full.data <- full.data %>% mutate(
  KSM_PROSPECT = case_when(
    KSM_PROSPECT_ACTIVE == TRUE ~ 'Current'
    , KSM_PROSPECT_ANY == TRUE ~ 'Past'
    , TRUE ~ 'No') %>% factor()
)
summary(full.data$KSM_PROSPECT)
boxplotter('KSM_PROSPECT')
```

Looks good.

```{r rv.cat.maxcashmo}
summary(full.data$GIVING_MAX_CASH_MO)
boxplotter('GIVING_MAX_CASH_MO')
```

Looks good.

### Numeric variables

Uses $\log_{10}$ transformation for dollars and $\sqrt{\ \ }$ transformation for counts.

```{r rv.num.ksmyr}
summary2(full.data$FIRST_KSM_YEAR)
scatterplotter('FIRST_KSM_YEAR', full.data %>% filter(FIRST_KSM_YEAR > 0))
```

Appears approximately linear until people graduating ~10 years ago.

```{r rv.num.gfirstyr}
summary2(full.data$GIVING_FIRST_YEAR)
scatterplotter('GIVING_FIRST_YEAR')
```

Same as previous.

```{r rv.num.gfirstcashamt}
summary2(full.data %>% select(GIVING_FIRST_YEAR_CASH_AMT) %>%
  filter(GIVING_FIRST_YEAR_CASH_AMT > 0) %>% unlist() %>% log10())
scatterplotter('GIVING_FIRST_YEAR_CASH_AMT'
               , full.data %>% filter(GIVING_FIRST_YEAR_CASH_AMT > 0)) +
  scale_x_log10(breaks = 10^(-2:7), labels = scales::dollar)
```

Nearly an $x = y$ dividing line; not expected to be exact given that `GIVING_MAX_CASH_AMT` is single gifts.
**CAUTION - endogenous**

```{r rv.num.gfirstplgamt}
summary2(full.data %>% select(GIVING_FIRST_YEAR_PLEDGE_AMT) %>%
  filter(GIVING_FIRST_YEAR_PLEDGE_AMT > 0) %>% unlist() %>% log10())
scatterplotter('GIVING_FIRST_YEAR_PLEDGE_AMT'
               , full.data %>% filter(GIVING_FIRST_YEAR_PLEDGE_AMT > 0)) +
  scale_x_log10(breaks = 10^(-2:7), labels = scales::dollar)
```

Looks ok, but probably need a cutoff.
**CAUTION - endogenous**

```{r rv.num.gct}
summary2(full.data %>% select(GIVING_CASH_TOTAL) %>%
  filter(GIVING_CASH_TOTAL > 0) %>% unlist() %>% log10())
scatterplotter('GIVING_CASH_TOTAL'
               , full.data %>% filter(GIVING_CASH_TOTAL > 0)) +
  scale_x_log10(breaks = 10^(-2:7), labels = scales::dollar)
```

Linear trend bounded by $y = x$ as expected.
**EXCLUDE**

```{r rv.num.gpt}
summary2(full.data %>% select(GIVING_PLEDGE_TOTAL) %>%
  filter(GIVING_PLEDGE_TOTAL > 0) %>% unlist() %>% log10())
scatterplotter('GIVING_PLEDGE_TOTAL'
               , full.data %>% filter(GIVING_PLEDGE_TOTAL > 0)) +
  scale_x_log10(breaks = 10^(-2:7), labels = scales::dollar)
```

Appears linear or piecewise linear, which follows given pledge payments.
**EXCLUDE**

```{r rv.num.gngct}
summary2(full.data %>% select(GIVING_NGC_TOTAL) %>%
  filter(GIVING_NGC_TOTAL > 0) %>% unlist() %>% log10())
scatterplotter('GIVING_NGC_TOTAL'
               , full.data %>% filter(GIVING_NGC_TOTAL > 0)) +
  scale_x_log10(breaks = 10^(-2:7), labels = scales::dollar)
```

Same as previous.
**EXCLUDE**

```{r rv.num.gallocs}
summary2(full.data$GIFTS_ALLOCS_SUPPORTED %>% sqrt())
scatterplotter('GIFTS_ALLOCS_SUPPORTED') +
  scale_x_sqrt()
```

Looks ok.

```{r rv.num.gfys}
summary2(full.data$GIFTS_FYS_SUPPORTED %>% sqrt())
scatterplotter('GIFTS_FYS_SUPPORTED') +
  scale_x_sqrt()
```

Looks ok.

```{r rv.num.gcash}
summary2(full.data$GIFTS_CASH %>% sqrt())
scatterplotter('GIFTS_CASH') +
  scale_x_sqrt()
```

Would want to exclude the influential point around $x = 300$ if this is used.
**Consider additional transformation**

```{r rv.num.gccard}
summary2(full.data$GIFTS_CREDIT_CARD %>% sqrt())
scatterplotter('GIFTS_CREDIT_CARD') +
  scale_x_sqrt()
```

Looks ok; candidate for discretizing.
**Consider additional transformation**

```{r rv.num.gstock}
summary2(full.data$GIFTS_STOCK %>% sqrt())
scatterplotter('GIFTS_STOCK') +
  scale_x_sqrt()
```

Completely different 0 and 1+ slopes. Could be a candidate for binary indicator.
**Consider additional transformation**

```{r rv.num.cpfy1}
rbind(
    summary2(full.data %>% select(CASH_PFY1) %>% filter(CASH_PFY1 > 0) %>% unlist() %>% log10())
  , summary2(full.data %>% select(CASH_PFY2) %>% filter(CASH_PFY2 > 0) %>% unlist() %>% log10())
  , summary2(full.data %>% select(CASH_PFY3) %>% filter(CASH_PFY3 > 0) %>% unlist() %>% log10())
  , summary2(full.data %>% select(CASH_PFY4) %>% filter(CASH_PFY4 > 0) %>% unlist() %>% log10())
  , summary2(full.data %>% select(CASH_PFY5) %>% filter(CASH_PFY5 > 0) %>% unlist() %>% log10())
)

# Grid plot
grid.arrange(
    scatterplotter('CASH_PFY1', full.data %>% filter(CASH_PFY1 > 0)) +
      scale_x_log10(breaks = 10^(-2:7), labels = scales::dollar)
  , scatterplotter('CASH_PFY2', full.data %>% filter(CASH_PFY2 > 0)) +
      scale_x_log10(breaks = 10^(-2:7), labels = scales::dollar)
  , scatterplotter('CASH_PFY3', full.data %>% filter(CASH_PFY3 > 0)) +
      scale_x_log10(breaks = 10^(-2:7), labels = scales::dollar)
  , scatterplotter('CASH_PFY4', full.data %>% filter(CASH_PFY4 > 0)) +
      scale_x_log10(breaks = 10^(-2:7), labels = scales::dollar)
  , scatterplotter('CASH_PFY5', full.data %>% filter(CASH_PFY5 > 0)) +
      scale_x_log10(breaks = 10^(-2:7), labels = scales::dollar)
)
```

High year-to-year correlations as expected.
**CAUTION - endogenous**

```{r rv.num.ngcpfy1}
rbind(
    summary2(full.data %>% select(NGC_PFY1) %>% filter(NGC_PFY1 > 0) %>% unlist() %>% log10())
  , summary2(full.data %>% select(NGC_PFY2) %>% filter(NGC_PFY2 > 0) %>% unlist() %>% log10())
  , summary2(full.data %>% select(NGC_PFY3) %>% filter(NGC_PFY3 > 0) %>% unlist() %>% log10())
  , summary2(full.data %>% select(NGC_PFY4) %>% filter(NGC_PFY4 > 0) %>% unlist() %>% log10())
  , summary2(full.data %>% select(NGC_PFY5) %>% filter(NGC_PFY5 > 0) %>% unlist() %>% log10())
)

# Grid plot
grid.arrange(
    scatterplotter('NGC_PFY1', full.data %>% filter(NGC_PFY1 > 0)) +
      scale_x_log10(breaks = 10^(-2:7), labels = scales::dollar)
  , scatterplotter('NGC_PFY2', full.data %>% filter(NGC_PFY2 > 0)) +
      scale_x_log10(breaks = 10^(-2:7), labels = scales::dollar)
  , scatterplotter('NGC_PFY3', full.data %>% filter(NGC_PFY3 > 0)) +
      scale_x_log10(breaks = 10^(-2:7), labels = scales::dollar)
  , scatterplotter('NGC_PFY4', full.data %>% filter(NGC_PFY4 > 0)) +
      scale_x_log10(breaks = 10^(-2:7), labels = scales::dollar)
  , scatterplotter('NGC_PFY5', full.data %>% filter(NGC_PFY5 > 0)) +
      scale_x_log10(breaks = 10^(-2:7), labels = scales::dollar)
)
```

Same as above.
**CAUTION - endogenous**

```{r rv.num.klc}
summary2(full.data$GIFT_CLUB_KLC_YRS %>% sqrt())
scatterplotter('GIFT_CLUB_KLC_YRS') +
  scale_x_sqrt()
```

Surprisingly close to linear; looks ok.

```{r rv.num.beq}
summary2(full.data$GIFT_CLUB_BEQUEST_YRS %>% sqrt())
scatterplotter('GIFT_CLUB_BEQUEST_YRS') +
  scale_x_sqrt()
```

Only 1 point with 2, so good candidate for binary indicator.
**Consider additional transformation**

```{r rv.num.loy}
summary2(full.data$GIFT_CLUB_LOYAL_YRS %>% sqrt())
scatterplotter('GIFT_CLUB_LOYAL_YRS') +
  scale_x_sqrt()
```

Looks ok but check how many are at each level.
**Consider additional transformation**

```{r rv.num.nulc}
summary2(full.data$GIFT_CLUB_NU_LDR_YRS %>% sqrt())
scatterplotter('GIFT_CLUB_NU_LDR_YRS') +
  scale_x_sqrt()
```

Looks ok but check how many are at each level.
**Consider additional transformation**

```{r rv.num.visits}
rbind(
    summary2(full.data$VISITS_PFY1 %>% sqrt())
  , summary2(full.data$VISITS_PFY2 %>% sqrt())
  , summary2(full.data$VISITS_PFY3 %>% sqrt())
  , summary2(full.data$VISITS_PFY4 %>% sqrt())
  , summary2(full.data$VISITS_PFY5 %>% sqrt())
)

# Grid plot
grid.arrange(
    scatterplotter('VISITS_PFY1') + scale_x_sqrt()
  , scatterplotter('VISITS_PFY2') + scale_x_sqrt()
  , scatterplotter('VISITS_PFY3') + scale_x_sqrt()
  , scatterplotter('VISITS_PFY4') + scale_x_sqrt()
  , scatterplotter('VISITS_PFY5') + scale_x_sqrt()
)
```

Hard to see what's going on past 6 visits; seems to be about linear until then in each case. Try combining 5 years of visit counts.

```{r rv.num.totvisits}
full.data <- full.data %>% mutate(
  VISITS = VISITS_PFY1 + VISITS_PFY2 + VISITS_PFY3 + VISITS_PFY4 + VISITS_PFY5
)
summary2(full.data$VISITS)
scatterplotter('VISITS') + scale_x_sqrt()
```

More of a logarithmic relationship.
**Consider log transformation**

```{r rv.num.visitors}
rbind(
    summary2(full.data$VISITORS_PFY1 %>% sqrt())
  , summary2(full.data$VISITORS_PFY2 %>% sqrt())
  , summary2(full.data$VISITORS_PFY3 %>% sqrt())
  , summary2(full.data$VISITORS_PFY4 %>% sqrt())
  , summary2(full.data$VISITORS_PFY5 %>% sqrt())
)

# Grid plot
grid.arrange(
    scatterplotter('VISITORS_PFY1') + scale_x_sqrt()
  , scatterplotter('VISITORS_PFY2') + scale_x_sqrt()
  , scatterplotter('VISITORS_PFY3') + scale_x_sqrt()
  , scatterplotter('VISITORS_PFY4') + scale_x_sqrt()
  , scatterplotter('VISITORS_PFY5') + scale_x_sqrt()
)
```

Try combining 5 years of visitor counts.

```{r rv.num.totvisitors}
full.data <- full.data %>% mutate(
  VISITORS = VISITORS_PFY1 + VISITORS_PFY2 + VISITORS_PFY3 + VISITORS_PFY4 + VISITORS_PFY5
)
summary2(full.data$VISITORS)
scatterplotter('VISITORS') + scale_x_sqrt()
```

Looks ok.

```{r rv.num.cmteenud}
summary2(full.data$COMMITTEE_NU_DISTINCT %>% sqrt())
scatterplotter('COMMITTEE_NU_DISTINCT') +
  scale_x_sqrt()
```

Looks like it might be piecewise-linear; consider discretizing.
**Consider additional transformation**

```{r rv.num.cmteenua}
summary2(full.data$COMMITTEE_NU_ACTIVE %>% sqrt())
scatterplotter('COMMITTEE_NU_ACTIVE') +
  scale_x_sqrt()
```

Likely unhelpful since I don't know committees the year largest gift was made.
**EXCLUDE**

```{r rv.num.cmteenuyr}
summary2(full.data$COMMITTEE_NU_YEARS %>% sqrt())
scatterplotter('COMMITTEE_NU_YEARS') +
  scale_x_sqrt()
```

Looks fine.

```{r rv.num.cmteeksmd}
summary2(full.data$COMMITTEE_KSM_DISTINCT %>% sqrt())
scatterplotter('COMMITTEE_KSM_DISTINCT') +
  scale_x_sqrt()
```

Can probably get away with linear, or discretize.
**Consider additional transformation**

```{r rv.num.cmteeksma}
summary2(full.data$COMMITTEE_KSM_ACTIVE %>% sqrt())
scatterplotter('COMMITTEE_KSM_ACTIVE') +
  scale_x_sqrt()
```

Likely unhelpful since I don't know committees the year largest gift was made.
**EXCLUDE**

```{r rv.num.cmteeksmyr}
summary2(full.data$COMMITTEE_KSM_YEARS %>% sqrt())
scatterplotter('COMMITTEE_KSM_YEARS') +
  scale_x_sqrt()
```

Looks fine.

```{r rv.num.cmteeksmldr}
summary2(full.data$COMMITTEE_KSM_LDR %>% sqrt())
scatterplotter('COMMITTEE_KSM_LDR') +
  scale_x_sqrt()
```

Looks ok but consider combining based on counts.
**Consider additional transformation**

```{r rv.num.cmteeksmldra}
summary2(full.data$COMMITTEE_KSM_LDR_ACTIVE %>% sqrt())
scatterplotter('COMMITTEE_KSM_LDR_ACTIVE') +
  scale_x_sqrt()
```

Nonmonontonic relationship; does that make sense? Artifact?
**EXCLUDE**

```{r rv.num.eventa}
summary2(full.data$KSM_EVENTS_ATTENDED %>% sqrt())
scatterplotter('KSM_EVENTS_ATTENDED') +
  scale_x_sqrt()
```

Looks ok.

```{r rv.num.eventyr}
summary2(full.data$KSM_EVENTS_YRS %>% sqrt())
scatterplotter('KSM_EVENTS_YRS') +
  scale_x_sqrt()
```

Looks ok.

```{r rv.num.event3fy}
summary2(full.data$KSM_EVENTS_PREV_3_FY %>% sqrt())
scatterplotter('KSM_EVENTS_PREV_3_FY') +
  scale_x_sqrt()
```

Looks ok.

```{r rv.num.eventreun}
summary2(full.data$KSM_EVENTS_REUNIONS %>% sqrt())
scatterplotter('KSM_EVENTS_REUNIONS') +
  scale_x_sqrt()
```

Looks ok, but could also work as an indicator; how many people have really gone to 4+?
**Consider additional transformation**

```{r rv.num.spkyrs}
summary2(full.data$KSM_SPEAKER_YEARS %>% sqrt())
scatterplotter('KSM_SPEAKER_YEARS') +
  scale_x_sqrt()
```

Just started tracking. Good candidate for binary indicator.
**Consider additional transformation**

```{r rv.num.spktimes}
summary2(full.data$KSM_SPEAKER_TIMES %>% sqrt())
summary(full.data$KSM_SPEAKER_TIMES %>% factor())
scatterplotter('KSM_SPEAKER_TIMES') +
  scale_x_sqrt()
```

Good candidate for binary indicator, though that's duplicative with the above.
**Consider additional transformation**

```{r rv.num.featyrs}
summary2(full.data$KSM_FEATURED_COMM_YEARS %>% sqrt())
scatterplotter('KSM_FEATURED_COMM_YEARS') +
  scale_x_sqrt()
```

Good candidate for binary indicator.
**Consider additional transformation**

```{r rv.num.feattimes}
summary2(full.data$KSM_FEATURED_COMM_TIMES %>% sqrt())
summary(full.data$KSM_FEATURED_COMM_TIMES %>% factor())
scatterplotter('KSM_FEATURED_COMM_TIMES') +
  scale_x_sqrt()
```

Good candidate for binary indicator, though that's duplicative with the above.
**EXCLUDE**

```{r rv.num.corpyrs}
summary2(full.data$KSM_CORP_RECRUITER_YEARS %>% sqrt())
scatterplotter('KSM_CORP_RECRUITER_YEARS') +
  scale_x_sqrt()
```

Looks fine; treat as binary indicator.

```{r rv.num.corptimes}
summary2(full.data$KSM_CORP_RECRUITER_TIMES %>% sqrt())
summary(full.data$KSM_CORP_RECRUITER_TIMES %>% factor())
scatterplotter('KSM_CORP_RECRUITER_TIMES') +
  scale_x_sqrt()
```

Duplicative with above.
**EXCLUDE**

```{r rv.num.athyrs}
summary2(full.data$ATHLETICS_TICKET_YEARS %>% sqrt())
summary(full.data$ATHLETICS_TICKET_YEARS %>% factor())
scatterplotter('ATHLETICS_TICKET_YEARS') +
  scale_x_sqrt()
```

Looks fine.

```{r rv.num.athtimes}
summary2(full.data %>% filter(ATHLETICS_TICKET_LAST > 0) %>% select(ATHLETICS_TICKET_LAST) %>%
           unlist())
scatterplotter('ATHLETICS_TICKET_LAST'
               , data = full.data %>% filter(ATHLETICS_TICKET_LAST > 0)) +
  scale_x_continuous(breaks = 2000:2050)
```

Looks fine.

## Additional exploration

### Count of cash gifts

```{r rv2.gifts_cash_hist}
full.data %>% filter(GIFTS_CASH > 0) %>% select(GIFTS_CASH) %>%
  mutate(GIFTS_CASH = cut(GIFTS_CASH, 50)) %>% summary()

histogrammer('GIFTS_CASH', data = full.data %>% filter(GIFTS_CASH > 0), m = 50)
```

Looks like around 95% of the donors have given no more than 25 cash gifts.

```{r rv2.gifts_cash}
# Full dataset regression line for comparison
lmtmp <- full.data %>% filter(GIVING_MAX_CASH_AMT > 0) %>%
  lm(log10(GIVING_MAX_CASH_AMT) ~ sqrt(GIFTS_CASH), data = .)

# Plot
scatterplotter('GIFTS_CASH', data = full.data %>% filter(GIFTS_CASH < 25)) +
  # Filtered data
  geom_point(aes(x = GIFTS_CASH, y = GIVING_MAX_CASH_AMT),
             data = full.data %>% filter(GIFTS_CASH >= 25),
             color = 'gray', alpha = .5) +
  scale_x_sqrt() +
  geom_abline(slope = lmtmp$coefficients[2], intercept = lmtmp$coefficients[1],
              linetype = 'solid', color = 'purple', alpha = .5)

remove(lmtmp)
```

The red line is a linear regression for $x < 25$ (excluding gray points) while the purple line is for the full dataset. Turns out that the (relatively) extreme values really make no difference.

### Count of credit card gifts

```{r rv2.gifts_cc_hist}
full.data %>% filter(GIFTS_CREDIT_CARD > 0) %>% select(GIFTS_CREDIT_CARD) %>%
  mutate(GIFTS_CREDIT_CARD = cut(GIFTS_CREDIT_CARD, 25)) %>% summary()

histogrammer('GIFTS_CREDIT_CARD', data = full.data %>% filter(GIFTS_CREDIT_CARD > 0), m = 25)

quantile(full.data %>% filter(GIFTS_CREDIT_CARD > 0) %>% select(GIFTS_CREDIT_CARD) %>% unlist, .9)
```

90th quantile is at $x = 8$.

```{r rv2.gifts_cc}
full.data %>% filter(GIVING_MAX_CASH_AMT > 0) %>%
  mutate(
    DISCRETE_CC = cut(GIFTS_CREDIT_CARD, breaks = c(-1:9, 1E99))
  ) %>%
boxplotter('DISCRETE_CC', data = .) +
  stat_summary(fun.y = mean, geom = 'point', size = 1, color = 'blue')
```

Honestly seems like **discretizing into 0, 1, and 2+ would work fine.**
I suspect that the dip for very high CC gift count entities is due to recurring (monthly) transactions.

### Binary candidates

```{r rv2.gifts_stock}
full.data %>% mutate(DISCRETE_STOCK = cut(GIFTS_STOCK, c(-1:9, 1E99))) %>% 
boxplotter('DISCRETE_STOCK', data = .) +
  stat_summary(fun.y = mean, geom = 'point', size = 1, color = 'blue')
```

**Treat `GIFTS_STOCK` as binary.**

```{r rv2.gift_clubs_beq}
full.data %>% mutate(GIFT_CLUB_BEQUEST_YRS = cut(GIFT_CLUB_BEQUEST_YRS, c(-1:2))) %>% 
boxplotter('GIFT_CLUB_BEQUEST_YRS', data = .) +
  stat_summary(fun.y = mean, geom = 'point', size = 1, color = 'blue')
```

**Treat `GIFT_CLUB_BEQUEST_YRS` as binary.**

```{r rv2.speaker_yrs}
full.data %>% mutate(KSM_SPEAKER_YEARS = cut(KSM_SPEAKER_YEARS, c(-1:2, 1E99))) %>% 
boxplotter('KSM_SPEAKER_YEARS', data = .) +
  stat_summary(fun.y = mean, geom = 'point', size = 1, color = 'blue')
```

```{r rv2.speaker_times}
full.data %>% mutate(KSM_SPEAKER_TIMES = cut(KSM_SPEAKER_TIMES, c(-1:8, 1E99))) %>% 
boxplotter('KSM_SPEAKER_TIMES', data = .) +
  stat_summary(fun.y = mean, geom = 'point', size = 1, color = 'blue')
```

**Use binary indicator on `KSM_SPEAKER_YEARS` and exclude `KSM_SPEAKER_TIMES`.**

### KSM leadership committees

```{r rv2.ksm_ldr_com}
full.data %>% mutate(COMMITTEE_KSM_LDR = cut(COMMITTEE_KSM_LDR, c(-1:6, 1E99))) %>% 
boxplotter('COMMITTEE_KSM_LDR', data = .) +
  stat_summary(fun.y = mean, geom = 'point', size = 1, color = 'blue')
```

Seems fine to leave as is.

### KSM Reunion attendance

```{r rv2.reunion_attendance}
full.data %>% mutate(KSM_EVENTS_REUNIONS = cut(KSM_EVENTS_REUNIONS, c(-1:6, 1E99))) %>% 
boxplotter('KSM_EVENTS_REUNIONS', data = .) +
  stat_summary(fun.y = mean, geom = 'point', size = 1, color = 'blue')
```

**Split `KSM_EVENTS_REUNIONS` into 0, 1, 2, 3+.**

### Featured communication years count

```{r rv2.featured_comms}
full.data %>% mutate(KSM_FEATURED_COMM_YEARS = cut(KSM_FEATURED_COMM_YEARS, c(-1:3, 1E99))) %>% 
boxplotter('KSM_FEATURED_COMM_YEARS', data = .) +
  stat_summary(fun.y = mean, geom = 'point', size = 1, color = 'blue')
```

This is somewhat problematic because people are often featured because they gave. Might as well try it though.
**Treat `KSM_FEATURED_COMM_YEARS` as binary.**

### Loyal gift club

```{r rv2.gc_loyal}
full.data %>% mutate(GIFT_CLUB_LOYAL_YRS = cut(GIFT_CLUB_LOYAL_YRS, c(-1:6, 1E99))) %>%
  select(GIFT_CLUB_LOYAL_YRS) %>% summary()

full.data %>% mutate(GIFT_CLUB_LOYAL_YRS = cut(GIFT_CLUB_LOYAL_YRS, c(-1:6, 1E99))) %>% 
boxplotter('GIFT_CLUB_LOYAL_YRS', data = .) +
  stat_summary(fun.y = mean, geom = 'point', size = 1, color = 'blue')
```

Looks fine as-is; good proportion within each bin.

### Non-KSM NU leadership gift clubs

```{r rv2.gc_nu_ldr}
full.data %>% mutate(GIFT_CLUB_NU_LDR_YRS = cut(GIFT_CLUB_NU_LDR_YRS, c(-1:16, 1E99))) %>%
  select(GIFT_CLUB_NU_LDR_YRS) %>% table()

full.data %>% mutate(GIFT_CLUB_NU_LDR_YRS = cut(GIFT_CLUB_NU_LDR_YRS, c(-1:16, 1E99))) %>% 
boxplotter('GIFT_CLUB_NU_LDR_YRS', data = .) +
  stat_summary(fun.y = mean, geom = 'point', size = 1, color = 'blue')
```

Practically no difference between the 1+ groups; appears fine to treat as binary.
**Treat `GIFT_CLUB_NU_LDR_YRS` as binary.**

## Derived features

### Visits

Does total visits (in the last 5 FY) make sense as a sqrt or log10 variable?

```{r rv2.visits}
scatterplotter('VISITS') + scale_x_sqrt()
scatterplotter('VISITS') + scale_x_log10()
```

Much more convincing on a log-10 scale.
**Use summed `VISITS` on a log-10 scale.**

### X out of 5 loyalty

Does `GIVING_MAX_CASH_AMT` vary with number of years giving in last 5?

Let $FY_t$ denote fiscal year cash giving in year $t$. If $\mathbf{1}$ is the indicator function:

$$ P_5 = \frac{1}{5} \left[ \sum_{i=0}^{4} \mathbf{1}\left\{ FY_{t-i} > 0 \right\} \right] $$

```{r rv2.x5_loyalty}
full.data %>% mutate(
  LAST_5_PCT = 1/5 * (
    ifelse(CASH_PFY1 > 0, 1, 0)
    + ifelse(CASH_PFY2 > 0, 1, 0)
    + ifelse(CASH_PFY3 > 0, 1, 0)
    + ifelse(CASH_PFY4 > 0, 1, 0)
    + ifelse(CASH_PFY5 > 0, 1, 0)
  )
) %>% select(LAST_5_PCT) %>% table()

full.data %>% mutate(
  LAST_5_PCT = {1/5 * (
    ifelse(CASH_PFY1 > 0, 1, 0)
    + ifelse(CASH_PFY2 > 0, 1, 0)
    + ifelse(CASH_PFY3 > 0, 1, 0)
    + ifelse(CASH_PFY4 > 0, 1, 0)
    + ifelse(CASH_PFY5 > 0, 1, 0)
  )} %>% factor()
) %>% boxplotter('LAST_5_PCT', data = .) +
  stat_summary(fun.y = mean, geom = 'point', size = 1, color = 'blue')
```

Very nice linear trend.
**Include `LAST_5_PCT` as a variable.**

### Upgrade behavior

Does `GIVING_MAX_CASH_AMT` vary as a function of upgrades/downgrades?

$$ U_5 = \sum_{i=0}^{4} \text{sign}\left( FY_{t-i} - FY_{t-i-1} \right) $$

```{r rv2.upgrade_5}
full.data %>% mutate(
  UPGRADE5 =
      sign(CASH_PFY1 - CASH_PFY2)
    + sign(CASH_PFY2 - CASH_PFY3)
    + sign(CASH_PFY3 - CASH_PFY4)
    + sign(CASH_PFY4 - CASH_PFY5)
) %>% select(UPGRADE5) %>% table()

full.data %>% mutate(
  UPGRADE5 = factor(
      sign(CASH_PFY1 - CASH_PFY2)
    + sign(CASH_PFY2 - CASH_PFY3)
    + sign(CASH_PFY3 - CASH_PFY4)
    + sign(CASH_PFY4 - CASH_PFY5)
  )
) %>% boxplotter('UPGRADE5', data = .) +
  stat_summary(fun.y = mean, geom = 'point', size = 1, color = 'blue')
```

Clearly nonlinear. Based on the counts, probably treat this as -2 to +2.

Try 3 years:

```{r rv2.upgrade_3}
full.data %>% mutate(
  UPGRADE3 =
      sign(CASH_PFY1 - CASH_PFY2)
    + sign(CASH_PFY2 - CASH_PFY3)
) %>% select(UPGRADE3) %>% table()

full.data %>% mutate(
  UPGRADE3 = factor(
      sign(CASH_PFY1 - CASH_PFY2)
    + sign(CASH_PFY2 - CASH_PFY3)
  )
) %>% boxplotter('UPGRADE3', data = .) +
  stat_summary(fun.y = mean, geom = 'point', size = 1, color = 'blue')
```

This is rather counterintuitive: people who have downgraded are more likely to have given a large gift compared to constant donors? Is this a Reunion effect?
**Include `UPGRADE3` giving behaviors.**

### Giving velocity

I'll use the [Blackbaud](https://www.kb.blackbaud.com/articles/Article/57194) definition, which compares last to three previous years of giving.

$$ V_3 = \left\{
  \begin{array}{l l}
    \frac{FY_t}{\frac{1}{3} \sum_{i=1}^{3}FY_{t-i}} \text{  when  } \sum_{i=1}^{3}FY_{t-i} > 0 \\
    1 \text{  otherwise}
  \end{array}
  \right. $$

```{r rv2.velocity_3}
full.data %>% mutate(
    denom = 1/3 * (CASH_PFY2 + CASH_PFY3 + CASH_PFY4)
  , VELOCITY3 = CASH_PFY1 / ifelse(denom == 0, max(CASH_PFY1, 1E-99), denom)
) %>% histogrammer('VELOCITY3', data = ., m = 50) +
  scale_x_log10(breaks = 10^(-5:5))

full.data %>% mutate(
    denom = 1/3 * (CASH_PFY2 + CASH_PFY3 + CASH_PFY4)
  , VELOCITY3 = CASH_PFY1 / ifelse(denom == 0, max(CASH_PFY1, 1E-99), denom)
) %>% filter(GIVING_MAX_CASH_AMT > 0) %>%
  scatterplotter('VELOCITY3', data = .) +
    scale_x_log10(breaks = 10^(-5:5))
```

Well, that's unexpected. There are likely interactions with age and Reunion cycle. Evidence for 4-6 clusters.

  * Nondonors during the entire period
  * V3 = -Inf (?)
  * 1E-8 < V3 < 1E-2
  * 1E-2 < V3 < 1
  * V3 = 1 (?)
  * V3 > 1

```{r rv2.velocity_discrete}
full.data %>% mutate(
    denom = 1/3 * (CASH_PFY2 + CASH_PFY3 + CASH_PFY4)
  , VELOCITY3 = CASH_PFY1 / ifelse(denom == 0, max(CASH_PFY1, 1E-99), denom)
  , VELOCITY_DISCR = case_when(
      CASH_PFY1 == 0 & denom == 0 ~ 'A. Non'
    , VELOCITY3 < 1E-8 ~ 'B. Way down'
    , VELOCITY3 < 1E-2 ~ 'C. Moderate down'
    , VELOCITY3 < 1 ~ 'D. Slight down'
    , VELOCITY3 == 1 ~ 'E. Unchanged'
    , VELOCITY3 > 1 ~ 'F. Up'
  )
) %>% select(VELOCITY_DISCR) %>% table()

full.data %>% mutate(
    denom = 1/3 * (CASH_PFY2 + CASH_PFY3 + CASH_PFY4)
  , VELOCITY3 = CASH_PFY1 / ifelse(denom == 0, max(CASH_PFY1, 1E-99), denom)
  , VELOCITY_DISCR = case_when(
      CASH_PFY1 == 0 & denom == 0 ~ 'A. Non'
    , VELOCITY3 < 1E-8 ~ 'B. Way down'
    , VELOCITY3 < 1E-2 ~ 'C. Moderate down'
    , VELOCITY3 < 1 ~ 'D. Slight down'
    , VELOCITY3 == 1 ~ 'E. Unchanged'
    , VELOCITY3 > 1 ~ 'F. Up'
  )
) %>% boxplotter('VELOCITY_DISCR', data = .) +
  stat_summary(fun.y = mean, geom = 'point', size = 1, color = 'blue')
```

Not super-clean but might as well see if it has any predictive power.
**Include `VELOCITY3`.**
**Include `VELOCITY_DISCR` as a factor.**

Alternate velocity definition: take *difference* between last year and average giving, rather than the ratio?

```{r rv2.velocity2_3}
full.data <- full.data %>% mutate(
  avg = 1/3 * (CASH_PFY2 + CASH_PFY3 + CASH_PFY4)
  , VELOCITY3B = CASH_PFY1 - avg
)

full.data %>%
  histogrammer('VELOCITY3B', data = ., m = 50) +
  scale_y_sqrt()

full.data %>% filter(GIVING_MAX_CASH_AMT > 0) %>%
  scatterplotter('VELOCITY3B', data = .)

grid.arrange(
  full.data %>% filter(GIVING_MAX_CASH_AMT > 0, VELOCITY3B > 0) %>%
    ggplot(aes(x = VELOCITY3B, y = GIVING_MAX_CASH_AMT)) +
    geom_point(alpha = .6, color = 'darkgreen') +
    geom_smooth() +
    geom_smooth(method = 'lm', color = 'green') +
    scale_y_log10(breaks = 20^(-2:7), labels = scales::dollar) +
    scale_x_log10()
, full.data %>% filter(GIVING_MAX_CASH_AMT > 0, VELOCITY3B < 0) %>%
    ggplot(aes(x = abs(VELOCITY3B), y = GIVING_MAX_CASH_AMT)) +
    geom_point(alpha = .6, color = 'darkred') +
    geom_smooth() +
    geom_smooth(method = 'lm', color = 'red') +
    scale_y_log10(breaks = 20^(-2:7), labels = scales::dollar) +
    scale_x_log10()
)

# Log10 scale preserving sign
full.data %>% mutate(VELOCITY3B_SGN = log10(abs(VELOCITY3B) + 1) * sign(VELOCITY3B)) %>%
  scatterplotter('VELOCITY3B_SGN', data = .)

full.data %>% mutate(VELOCITY3B_SGN_DOUBLING = {log2(abs(VELOCITY3B) + 1) * sign(VELOCITY3B)} %>%
                       round() %>% factor()) %>%
  boxplotter('VELOCITY3B_SGN_DOUBLING', data = .) +
  stat_summary(fun.y = mean, geom = 'point', size = 1, color = 'blue')
```

Wonder why Blackbaud doesn't use this? So much cleaner.
**Include `VELOCITY3B_SGN` as numeric.** Consider truncating to e.g. $\pm 4$

### Age imputation

Issue: various entities are either missing age or have something unreasonable (2 years old when receiving their MBA).

```{r age_nas}
full.data %>% select(DOB) %>% is.na() %>% sum()
```

Over 27,000 unknown age.

```{r age_impossible_DOB}
full.data %>% mutate(under_21 = FIRST_KSM_YEAR - year(DOB) < 21) %>%
  filter(year(DOB) > 1000 & FIRST_KSM_YEAR > 1000) %>% select(under_21) %>%
  unlist() %>% na.omit() %>% sum()

full.data %>% mutate(DOB_yr = year(DOB)) %>% filter(DOB_yr > 1000 & FIRST_KSM_YEAR > 1000) %>%
  ggplot(aes(x = DOB_yr, y = FIRST_KSM_YEAR)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 21, color = 'green') +
  geom_abline(slope = 1, intercept = 15, color = 'yellow') +
  geom_abline(slope = 1, intercept = 0, color = 'red')
```

Between green and yellow lines are supposedly people < 21 when graduating; yellow and red are people < 15; under red line means graduation year is before they were born.

Two possibilities: age imputation, or use a comparable but more accurate metric.

Proposal: record year, defined as `FIRST_KSM_YEAR` for graduates, first year of giving for nongraduates if available, else record creation date. All grad years < 1908 can be set to 1908 (founding year).

# Data file cleanup goals

  * R script to clean data file for processing
  * Written as a function taking data file path as an argument
  
This is implemented in [parse_data.R](https://github.com/phively/ksm-models/blob/master/af-10k-fy17/scripts/parse_data.R).

