---
title: "03b AF 10K Ever Gave ROC"
output: html_notebook
---

# Goal

Produce ROC plots for the previous models.

This code may be run immediately after the previous [cross-validation steps](https://github.com/phively/ksm-models/blob/master/af-10k-fy17/03%20AF%2010K%20Ever%20Gave%20Classification.Rmd) have completed.

# ROC plots

Receiver operating characteristic (ROC) plots are a standard method used to assess classifiers. The greater the area under the curve the better. I'll check against the holdout data.

```{r roc.mat.gen}
roc_matrix_gen <- function(model, data) {
  data.frame(
      X1 = predict(model, newdata = data, type = 'response')
    , y = data$GAVE_10K
  ) %>% arrange(desc(X1)) %>%
  mutate(
      TPR = cumsum(y) / sum(y)
    , FPR = cumsum(y == 0) / sum(y == 0)
  ) %>%
  return()
}
```

```{r roc.plots}
rocdat <- cbind(model = 'baseline', roc_matrix_gen(mbaseline, holdoutdat)) %>%
  rbind(cbind(model = 'trimmed', roc_matrix_gen(mtrim3, holdoutdat))) %>%
  rbind(cbind(model = 'nonpen', roc_matrix_gen(mnpen, holdoutdat))) %>%
  rbind(cbind(model = 'pen', roc_matrix_gen(mpen, holdoutdat)))

rocdat %>%
  ggplot(aes(x = FPR, y = TPR, color = model)) +
  geom_line(size = 1) +
  geom_abline(slope = 1, intercept = 0, linetype = 'dashed', col = 'gray') +
  scale_x_continuous(breaks = seq(0, 1, by = .1), expand = c(0, 0)) +
  scale_y_continuous(breaks = seq(0, 1, by = .1), expand = c(0, 0)) +
  coord_equal() +
  labs(title = 'ROC plot')
```

```{r roc.auc}
# Riemann sum AUC
data.frame(
    baseline = with(
        rocdat %>% filter(model == 'baseline')
      , sum(1/nrow(holdoutdat) * TPR)
    )
  , trimmed = with(
        rocdat %>% filter(model == 'trimmed')
      , sum(1/nrow(holdoutdat) * TPR)
    )
  , nonpen = with(
        rocdat %>% filter(model == 'nonpen')
      , sum(1/nrow(holdoutdat) * TPR)
    )
  , pen = with(
        rocdat %>% filter(model == 'pen')
      , sum(1/nrow(holdoutdat) * TPR)
    )
)
```

No discernible difference between the models.