---
title: "04 AF baseline penalized splines model"
output: html_notebook
---

# Goal

This code reproducibly recreates the best model identified during the previous [cross-validation steps](https://github.com/phively/ksm-models/blob/master/af-10k-fy17/03%20AF%2010K%20Ever%20Gave%20Classification.Rmd).

# Setup

```{r load_packages}
source('scripts/load_packages.R')
source('scripts/parse_data.R')
```

Generate modeling data.

```{r gen_data}
full <- parse_data('data/2017-12-21 AF 10K Model.csv')
mdat <- full
source('scripts/parse_modeling_data.R')
```

# Modeling

Build the baseline penalized splines model using glmnet.

```{r model}
# Parallelize to ncores - 1
registerDoParallel(detectCores() - 1)

mbaseline <- cv.glmnet(as.factor(GAVE_10K) ~
                           ns(RECORD_YR, df = 5)
                         + PROGRAM_GROUP + PREF_ADDR_TYPE_CODE + HOUSEHOLD_CONTINENT
                         + GIVING_PLEDGE_ANY
                         + GIVING_PLEDGE_FIRST_YR + GIFTS_ALLOCS_SUPPORTED
                         + GIFTS_CREDIT_CARD + GIFTS_STOCK + GIFT_CLUB_KLC_YRS
                         + GIFT_CLUB_NU_LDR_YRS + GIFT_CLUB_LOYAL_YRS
                         + ns(VELOCITY3_LIN, df = 4)
                         , data = mdat
                         , family = 'binomial'
                         , alpha = 0
                         , lambda = NULL # Not specified; try a variety
                         , parallel = TRUE
)
```

Create predictions.

```{r preds}
preds <- predict(mbaseline, newdata = mdat, type = 'response', s = 'lambda.1se') %>%
  data.frame() %>%
  mutate(fitted = X1) %>%
  select(fitted)
```

Verify calibration results still hold.

```{r calibration}
# Penalized data frame
data.frame(
    class = mdat$GAVE_10K
  , prediction = preds$fitted
) %>%
# Plot results
ggplot(aes(x = prediction, y = class)) +
  geom_point(aes(color = as.factor(class))) +
  geom_smooth(method = 'gam', formula = y ~ s(x, k = 5)) +
  geom_abline(slope = 1, intercept = 0) +
  labs(title = 'GAM smoother, penalized coefficients', color = 'class')
```

Looks fine.

```{r roc.dat}
# Predicted versus actual
y <- data.frame(yhat = preds$fitted, y = mdat$GAVE_10K) %>%
  arrange(desc(yhat)) %>%
  mutate(
      TPR = cumsum(y) / sum(y)
    , FPR = cumsum(y == 0) / sum(y == 0)
  )

y %>%
  ggplot(aes(x = FPR, y = TPR)) +
  geom_line(color = 'blue') +
  geom_abline(slope = 1, intercept = 0, linetype = 'dashed', col = 'gray') +
  scale_x_continuous(breaks = seq(0, 1, by = .1), expand = c(0, 0)) +
  scale_y_continuous(breaks = seq(0, 1, by = .1), expand = c(0, 0)) +
  coord_equal() +
  labs(title = 'ROC plot')
```

The area under the curve can be quickly approximated as a Riemann sum.

```{r}
# AUC (Riemann sum)
sum(1/nrow(y) * y$TPR)
```

This is an outstanding result.

# Save results

Save to csv.

```{r save}
write.csv(cbind(preds, full), 'data/results.csv', row.names = FALSE)
```
